---
title: "InfantProdStudy"
author: "Catherine Laing"
date: "April 16, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r Set up, include=FALSE}
library(forcats)
library(effsize)
library(rcompanion)
library(tidyverse)
library(stringi)

multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
  library(grid)
  
  # Make a list from the ... arguments and plotlist
  plots <- c(list(...), plotlist)
  
  numPlots = length(plots)
  
  # If layout is NULL, then use 'cols' to determine layout
  if (is.null(layout)) {
    # Make the panel
    # ncol: Number of columns of plots
    # nrow: Number of rows needed, calculated from # of cols
    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
                     ncol = cols, nrow = ceiling(numPlots/cols))
  }
  
  if (numPlots==1) {
    print(plots[[1]])
    
  } else {
    # Set up the page
    grid.newpage()
    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))
    
    # Make each plot, in the correct location
    for (i in 1:numPlots) {
      # Get the i,j matrix positions of the regions that contain this subplot
      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))
      
      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
                                      layout.pos.col = matchidx$col))
    }
  }
}
```

```{r data, include=FALSE}
source("ProductionData_randsubj.R")
source("Data_scrambling_randsubj.R")
```

# Methods

Descriptively, 24/44 infants had at least one VMS consonant; 9 had two VMS in repertoire (5 females) two infants (one female) had three. The most common VMS consonant was /p,b/, which featured in 16 infants’ repertoires. Ten infants had /t,d/ in repertoire, and eight had /k, ɡ/ (see S1 and ST1). 

```{r Establishing VMS groups & ST1}

vmstotal %>% filter(VMSgroup == "withVMS") %>% group_by(sex) %>% tally() #24
vmscount %>% filter(Group == "Multi") %>% group_by(subj, VMS, sex) %>% tally() #11
vmscount %>% filter(Group == "One") %>% group_by(subj, VMS, sex) %>% tally() #13


vmstotal %>% filter(VMSgroup == "noVMS") %>% group_by(sex) %>% tally() #20 (8)
vmstotal %>% filter(VMS == "p b") %>% group_by(sex) %>% tally() # 8 (6)
vmstotal %>% filter(VMS == "t d") %>% group_by(sex) %>% tally() # 4 (1)
vmstotal %>% filter(VMS == "k g") %>% group_by(sex) %>% tally() # 1 (0)
vmstotal %>% filter(VMS == "p b m") %>% group_by(sex) %>% tally() # 1 (0)
vmstotal %>% filter(VMS == "p b t d") %>% group_by(sex) %>% tally() # 1 (1)
vmstotal %>% filter(VMS == "p b k g") %>% group_by(sex) %>% tally() # 4(2)
vmstotal %>% filter(VMS == "t d k g") %>% group_by(sex) %>% tally() # 2(1)
vmstotal %>% filter(VMS == "t d n") %>% group_by(sex) %>% tally() # 1(1)
vmstotal %>% filter(VMS == "p b t d n") %>% group_by(sex) %>% tally() # 1(0)
vmstotal %>% filter(VMS == "p b t d k g") %>% group_by(sex) %>% tally() # 1 (1)
```

Every consonant that the infant produced in their hour-long video at month 10 (n=27) or 11 (n=17)

```{r Infant age in months; Table 1}

vmstotal %>% group_by(month) %>% tally()

vmstotal %>% group_by(VMSgroup) %>% filter(month == 10) %>% tally()

vmstotal %>% group_by(VMSgroup) %>% filter(month == 11) %>% tally()

```

51% (n=976) of CPs were preceded by audible caregiver speech in this timeframe. The majority of these (74%) included a concrete noun.

```{r Counting data for CGprompt}

CPdata %>% tally() # 1916 CPs

CPdata %>% filter(cgprompt != "[none]" & 
                    cgprompt != "[unclear]" & 
                    cgprompt != "[unavailable]" & 
                    cgprompt != "[imitation]" &
                    cgprompt != "[media]") %>% 
  tally() # 976

976/1916 # 51%

CPdata %>% filter(wordclass != "[none]") %>% tally() # 976
CPdata %>% filter(wordclass == "N") %>% tally() # 720
CPdata %>% filter(wordclass != "N" & wordclass != "[none]") %>% tally() # 720


720/976  # 74%

```
Attended objects accounted for 62% of our CP data (n=1,179 CPs). After removing instances when the infant was attending to an object but its identity or label was unclear (n=78), we were left with 1,101 data points for the attended object analysis. 

```{r Counting data for AttObj}

CPdata %>% filter(object != "[none]" & object != "[unclear]" & object != "[out of view]") %>% tally() # 1179

1179/1916 # 62%

CPdata %>% filter(object == "[toy]") %>% tally() # 78

1179-78

```

For both caregiver input and attended object, in around 5% of cases (40 caregiver prompts and 58 attended objects) this meant there was no codable consonant for a word (e.g. neither wow nor  whisk have a supra-glottal consonant with full closure in syllable-initial position).
```{r Counting other data for CGprompt}

CPdata %>% filter(!grepl("^\\[",cgprompt) & C1Prompt == "none") %>% tally()  # 40 instances of CGprompts with no supraglottal consonant

CPdata %>% filter(!grepl("^\\[", object) & C1Obj == "none") %>% tally()  # 58 instances of Objects with no supraglottal consonant


 1916 - (40+58) # 1818
 (40+58)/1916 # 5%
```

# Results

An initial analysis of infant age, sex, and maternal education showed that these variables did not account for significant variance in the number of CP types or tokens, in audio-recordings or in videos, and thus data are collapsed across these dimensions in all further analysis. See S3 for further details.
```{r Data Analysis Plan - confounding variables - Audio data}

# CP Tokens 

#shapiro.test(vmstotal$audiotokens) # Not normal; 
                                    # log(audiotokens+1) also not normal so stick with non-transformed data and Non-para tests

wilcox.test(audiotokens ~ month, vmstotal) # ns p=.718

wilcox.test(audiotokens ~ sex, vmstotal) # ns p=.541

cor.test(vmstotal$MOTedulevel, vmstotal$audiotokens, method = 'spearman') # ns p=.273

# CP types

#shapiro.test(CPtypes_audio$CPtypes) # not normal

wilcox.test(CPtypes ~ sex, CPtypes_audio) # ns, p=.440
wilcox.test(CPtypes ~ month, CPtypes_audio) #ns, p=.564

cor.test(CPtypes_audio$MOTedulevel, CPtypes_audio$CPtypes, method = 'spearman') # ns p=.243

```


```{r Data Analysis Plan - counfounding variables - Video data}

#shapiro.test(ProdData$CPtokens) # Not normal
#shapiro.test(ProdData$CPtypes) # Not normal

wilcox.test(CPtokens ~ sex, ProdData) # p=.655

wilcox.test(CPtypes ~ sex, ProdData) # p=.971

wilcox.test(CPtokens ~ month, ProdData) # p=.596

wilcox.test(CPtypes ~ month, ProdData) # p=.363

cor.test(ProdData$MOTedulevel, ProdData$CPtokens, method = 'spearman') # p=.193
cor.test(ProdData$MOTedulevel, ProdData$CPtypes, method = 'spearman') # p=.886
```

None of the infants were outliers in our dataset (defined as producing more/less than 3 standard deviations from the mean CP types/tokens).
```{r outliers}

ProdData %>% mutate(outlier = (CPtokens > 
                      (mean(CPtokens, na.rm=T) + 
                         3*(sd(CPtokens, na.rm=T))) |
                      CPtokens <
                      (mean(CPtokens, na.rm=T) - 
                         3*(sd(CPtokens, na.rm=T))))) %>%
  filter(outlier == TRUE)  # no outliers

ProdData %>% mutate(outlier = (CPtypes > 
                      (mean(CPtypes, na.rm=T) + 
                         3*(sd(CPtypes, na.rm=T))) |
                      CPtypes <
                      (mean(CPtypes, na.rm=T) - 
                         3*(sd(CPtypes, na.rm=T))))) %>%
  filter(outlier == TRUE)  # no outliers

```

# Infant production: Audio recordings

On average, infants produced 97 consonant tokens in the half-hour segment of the daylong audio-recordings we analyzed. Variability was high (Range=0=610; see Table 3); three infants (2 females) produced zero consonants. The most prominent consonant was /p,b/ (n=1716), followed by /t,d/ (n=1364) and /k, g/ (n=745). 
```{r Audio recordings summary}

vmscount %>% group_by(subj) %>% summarise(total = sum(n)) %>% ungroup() %>% summarise(meanprod = mean(total), #96.59
                                                                                      sdprod = sd(total),  #116.64
                                                                                      minprod = min(total), #0
                                                                                      maxprod = max(total)) #610

vmscount %>% group_by(subj) %>% summarise(total = sum(n)) %>% filter(total == 0) %>% tally() #3 infants with 0 productions

vmscount %>% group_by(Consonant) %>% summarise(total = sum(n))

```

```{r Table 3 Audio}

# Audio

vmstotal %>% summarise(meantyp = mean(audiotypes),  # 3.23
                       sdtyp = sd(audiotypes))      # 1.43
 
vmstotal %>% group_by(VMSgroup) %>%                 
  summarise(meantyp = mean(audiotypes),             # 2.3(1.42)
                       sdtyp = sd(audiotypes))      # 4(.88)

vmscount %>% filter(VMSgroup == "withVMS") %>% 
  group_by(subj) %>% 
  summarise(total = sum(n)) %>% 
  ungroup() %>%
  summarise(meanprod = mean(total),                 # 165.75
                          sdprod = sd(total))       # 119.51
 
vmscount %>% filter(VMSgroup == "noVMS") %>% 
  group_by(subj) %>% 
  summarise(total = sum(n)) %>% 
  ungroup() %>%
  summarise(meanprod = mean(total),                 # 13.6
                          sdprod = sd(total))       # 14.41
```

```{r Table 3 Video}

# Video

CPdata %>% group_by(subj) %>% 
  summarise(videotypes = n_distinct(ctype)) %>% ungroup() %>%
  summarise(Mvideotypes = mean(videotypes),                     # 4.34
            SDvideotypes = sd(videotypes))                      # 1.27

CPdata %>% group_by(subj) %>% 
  filter(VMSgroup == "withVMS") %>%
  summarise(videotypes = n_distinct(ctype)) %>% ungroup() %>%
  summarise(Mvideotypes = mean(videotypes),                     # 4.67
            SDvideotypes = sd(videotypes))                      # 1.31

CPdata %>% group_by(subj) %>% 
  filter(VMSgroup == "noVMS") %>%
  summarise(videotypes = n_distinct(ctype)) %>% ungroup() %>%
  summarise(Mvideotypes = mean(videotypes),                     # 3.95
            SDvideotypes = sd(videotypes))                      # 1.15

CPdata %>% group_by(subj) %>% tally() %>%
  ungroup() %>%
  summarise(Mvideotok = mean(n),                                # 43.55
            SDvideotok = sd(n))                                 # 34.13

CPdata %>% filter(VMSgroup == "withVMS") %>%
  group_by(subj) %>% tally() %>%
  ungroup() %>%
  summarise(Mvideotok = mean(n),                                # 60.13
            SDvideotok = sd(n))                                 # 34.59

CPdata %>% filter(VMSgroup == "noVMS") %>%
  group_by(subj) %>% tally() %>%
  ungroup() %>%
  summarise(Mvideotok = mean(n),                                # 23.65
            SDvideotok = sd(n))                                 # 20.59


# Input prompt match

table(is.na(ProdData$Prod.Prompt.match.PC)) # all 44 infants have caregiver prompts prior to some CPs

ProdData %>% summarize(Promptmatch = mean(Prod.Prompt.match.PC, na.rm=T),  # .48
                   SDPromptmatch = sd(Prod.Prompt.match.PC, na.rm=T))      # .23

ProdData %>% group_by(VMSgroup) %>% summarize(Promptmatch = mean(Prod.Prompt.match.PC, na.rm=T),   # .55 vs. .41
                   SDPromptmatch = sd(Prod.Prompt.match.PC, na.rm=T))                              # .17 vs/ .28

# Attended object match
table(is.na(ProdData$Prod.Obj.match.PC)) # only 43 infants attend to objects during CP

ProdData %>% summarize(Objmatch = mean(Prod.Obj.match.PC, na.rm=T),         # .38
                   SDObjmatch = sd(Prod.Obj.match.PC, na.rm=T))             # .27

ProdData %>% group_by(VMSgroup) %>% summarize(Objmatch = mean(Prod.Obj.match.PC, na.rm=T),         # .49 vs. .25
                   SDObjmatch = sd(Prod.Obj.match.PC, na.rm=T))                                    # .27 vs. .22

```

As expected, withVMS infants produced more babble in the audio recordings than noVMS infants, and this difference was significant for both types and tokens (Types: Est.Diff.=1.99, δ=.99, p<.001, 95% CI=[0.99, 2.99]; Tokens: Est.Diff=130, δ=.69, p<.001, 95% CI=[103, 155]; by Wilcoxon Rank Sum test; see Table 3 & Figure 1).  

```{r comparing production across groups in audio recordings}

#shapiro.test(vmstotal$audiotypes) # not normal

wilcox.test(subset(vmstotal, VMSgroup == "withVMS")$audiotypes, 
            subset(vmstotal, VMSgroup == "noVMS")$audiotypes, conf.int=T) # p<.001

cliff.delta(d = subset(vmstotal, VMSgroup == "withVMS")$audiotypes, 
     f = subset(vmstotal, VMSgroup == "noVMS")$audiotypes)   # .69 large

#shapiro.test(vmstotal$audiotokens) # not normal

wilcox.test(subset(vmstotal, VMSgroup == "withVMS")$audiotokens, 
            subset(vmstotal, VMSgroup == "noVMS")$audiotokens, conf.int=T) # p<.001

cliff.delta(d = subset(vmstotal, VMSgroup == "withVMS")$audiotokens, 
     f = subset(vmstotal, VMSgroup == "noVMS")$audiotokens)   # .99 large

```

```{r Figure 1}

Figure1a <- ggplot(subset(vmscount, VMSgroup == "withVMS"), 
                   aes(x=Consonant, y=log10(n))) +
  stat_summary(fun.y=mean, geom = "point", 
               aes(group = subj, colour = Consonant), shape=1, size=2, 
               position = position_jitter(width = .1, height=0)) +
  stat_summary(fun.data=mean_cl_boot, geom = "pointrange", shape=2, size=1) +
  scale_x_discrete(limits = c("p", "t", "k", "m", "n"),
                   labels = c("pb", "td", "kg", "m", "n")) +
  theme_bw(base_size=15) +
  scale_y_continuous(breaks=c(0, .301, 1, 1.699, 2.69897), 
                     labels=c(0, 2, 10, 50, 500)) +
  xlab("Consonant Type") +
  ylab("n") +
  theme(legend.position = "none",
        axis.title = element_text(size=22),
        axis.text = element_text(size=22),
        strip.text = element_text(face="bold", size=22,lineheight=7.0)) +
  facet_wrap(~VMSgroup)

Figure1b <- ggplot(subset(vmscount, VMSgroup == "noVMS"), 
                   aes(x=Consonant, y=log10(n))) +
  stat_summary(fun.y=mean, geom = "point", 
               aes(group = subj, colour = Consonant), shape=1, size=2,
               position = position_jitter(width = .1, height=0)) +
  stat_summary(fun.data=mean_cl_boot, geom = "pointrange", shape=2, size=1) +
  scale_x_discrete(limits = c("p", "t", "k", "m", "n"),
                   labels = c("pb", "td", "kg", "m", "n")) +
  theme_bw(base_size=15) +
  scale_y_continuous(breaks=c(0, .301, 1, 1.477, 1.699, 2), 
                     labels=c(0, 2, 10, 30, 50, 100)) +
  xlab("Consonant Type") +
  ylab("n") +
  theme(legend.position = "none",
        axis.title = element_text(size=22),
           axis.text = element_text(size=22),
        strip.text = element_text(face="bold", size=22,lineheight=7.0)) +
  facet_wrap(~VMSgroup)

multiplot(Figure1a, Figure1b, cols=2)
```
  
# Infant Production: Video recordings


On average, infants produced 44 CP tokens in the hour-long video recordings and 4.34 CP types. See Table 3 and Figure 2. Once again, withVMS infants produced significantly more tokens than noVMS infants (60.13 vs. 23.65), while the number of CP types was marginally higher for withVMS infants (4.67 vs. 3.95; see Table 3 and 4). 

```{r Effect of VMS group on CP types/tokens}

wilcox.test(subset(ProdData, VMSgroup == "withVMS")$CPtokens, 
            subset(ProdData, VMSgroup == "noVMS")$CPtokens, conf.int=T) # p<.001

cliff.delta(d = subset(ProdData, VMSgroup == "withVMS")$CPtokens, 
     f = subset(ProdData, VMSgroup == "noVMS")$CPtokens)   # .68 large

ProdData %>% group_by(VMSgroup) %>% summarise(meantok = mean(CPtokens))

wilcox.test(subset(ProdData, VMSgroup == "withVMS")$CPtypes, 
            subset(ProdData, VMSgroup == "noVMS")$CPtypes, conf.int=T) # p=.06

cliff.delta(d = subset(ProdData, VMSgroup == "withVMS")$CPtypes, 
            f = subset(ProdData, VMSgroup == "noVMS")$CPtypes)   # .32 small

ProdData %>% group_by(VMSgroup) %>% summarise(meantyp = mean(CPtypes))

```

withVMS infants produced significantly more consonants that were in their stable VMS repertoire (INREP consonants) relative to other consonants (OUTREP consonants), as assessed by comparing the total number of INREP tokens/total number of VMS consonant types vs. the total number of OUTREP tokens/number of non-VMS consonant types produced in the video data (INREP: mean tokens=23.3; OUTREP: mean tokens=8.7; Est.Diff.=9.6, r=.63, p=.002, 95% CI=[4.08, 24.12]; paired Wilcoxon Signed-Rank test). 
  
```{r calcuating number of in vs out VMS consonants in video recordings}

compareRecordings %>% filter(VMS.Prod.Cong == T) %>% summarise(meanOverall = mean(meanProd),
                                                               sdOverall = sd(meanProd))

compareRecordings %>% filter(VMS.Prod.Cong == F) %>% summarise(meanOverall = mean(meanProd),
                                                               sdOverall = sd(meanProd))

#shapiro.test(compareRecordings$meanProd) # not normal

wilcox.test(compareRecordings.spread$VMS.Prod.Cong, 
            compareRecordings.spread$VMS.Prod.Incong, paired=T, conf.int=T)

wilcoxonPairedR(x = compareRecordings$meanProd,
                g = compareRecordings$VMS.Prod.Cong) # r= .63

```


```{r Figure 2}
Figure2a <- ggplot(ProdData, aes(x=VMSgroup, y=CPtokens)) +
  geom_violin(aes(fill = VMSgroup), alpha=.3) +
  stat_summary(fun.y=mean, geom = "point", 
               aes(group = subj), shape=1, size=2.5, position=position_jitter(height = 0, width = 0.01)) +
  stat_summary(fun.data=mean_cl_boot, geom = "pointrange", colour='red', shape=17, size=1) +
  scale_x_discrete(limits = c("withVMS", "noVMS"), 
                   labels = c("withVMS", "noVMS")) +
  theme_bw(base_size=15) +
  theme(axis.text = element_text(size=18),
           axis.title = element_text(size=18),
           legend.position="none") +
  ylab("n CP tokens") +
  xlab('')

Figure2b <- ggplot(ProdData, aes(x=VMSgroup, y=CPtypes)) +
    geom_violin(aes(fill = VMSgroup), alpha=.3) +
  stat_summary(fun.y=mean, geom = "point", 
               aes(group = subj), shape=1, size=2.5, position=position_jitter(height = 0, width = .1)) +
  stat_summary(fun.data=mean_cl_boot, geom = "pointrange", colour='red', shape=17, size=1) +
  scale_x_discrete(limits = c("withVMS", "noVMS"), 
                   labels = c("withVMS", "noVMS")) +
  coord_cartesian(ylim=c(0, 8)) +
  scale_y_continuous(breaks=seq(0, 8, 2))  +
  ylab("n CP types") +
  xlab('') +
  theme_bw(base_size=15) +
  theme(legend.position="none",
        axis.title = element_text(size=18),
        axis.text = element_text(size=18))

multiplot(Figure2a, Figure2b, cols=2)
```


## Input prompts

Overall, 1451 CPs in the data were produced alongside either an input prompt or an attended object; 527 CPs were produced alongside both an input prompt and an attended object.  
```{r Overview of context and CPs}

CPdata %>% mutate(nostim = ifelse(C1Obj == "none" & C1Prompt == "none", T, F)) %>%
  filter(nostim == FALSE) %>% tally() # 1451

CPdata %>% mutate(bothstim = ifelse(C1Obj != "none" & C1Prompt != "none", T, F)) %>%
  filter(bothstim == TRUE) %>% tally() # 527

CPdata %>% filter(C1Prompt != "none") %>% tally() # 936
```




On average, 51% of the infants’ CPs were preceded by caregiver speech (SD=.18), and 48% of these matched the salient word in this preceding segment (e.g. mother says ‘ball’ and baby says ‘ba’; SD=.23). We ran a Wilcoxon Rank Sum test to determine whether the proportion of each infant’s congruent CPs differed by VMS group. This difference was  not significant (Est.Diff.=.11, δ=.28, p=.111, 95% CI=[-.03, .3]; see Table 4); withVMS infants’ CPs matched caregiver prompts 14% more often than noVMS infants’ did (.55(.17) vs. .41(.28)). 
```{r Input prompt overview}

ProdData %>% summarize(meanPrompt = mean(nPrompt/CPtokens, na.rm=T), #.51
                   sdPrompt = sd(nPrompt/CPtokens, na.rm=T), #.18
                   Promptmatch = mean(Prod.Prompt.match.PC, na.rm=T), #.5
                   SDPromptmatch = sd(Prod.Prompt.match.PC, na.rm=T)) #.23

#shapiro.test(ProdData$Prod.Prompt.match.PC) # p=.048

#t.test(Prod.Prompt.match.PC ~ VMSgroup, ProdData) # p=.057

wilcox.test(subset(ProdData, VMSgroup=="withVMS")$Prod.Prompt.match.PC, 
            subset(ProdData, VMSgroup=="noVMS")$Prod.Prompt.match.PC, conf.int=T) # p=.11, EstDiff=-.11

cliff.delta(d = subset(ProdData, VMSgroup=="withVMS")$Prod.Prompt.match.PC, 
            f = subset(ProdData, VMSgroup=="noVMS")$Prod.Prompt.match.PC)   # .28 small

ProdData %>% group_by(VMSgroup) %>% summarise(meanInput = mean(Prod.Prompt.match.PC),
                                              sdInput = sd(Prod.Prompt.match.PC))

```

WithVMS infants’ CPs matched the consonantal properties of their caregivers’ speech significantly more in the real data set than in the scrambled one (.55(.17) vs. .22(.14), Est.Diff.=.33, δ=.90, p<.001, 95% CI=[.23, .43], by Wilcoxon test), as did noVMS infants’ CPs (.41(.28) vs. .2(.16), Est.Diff.=.24, δ=.45, p=.014, 95% CI=[.04, .37]). 
```{r Input prompt data vs scrambled data}

prompt.real <- ProdData %>%
  select(subj, month, VMSgroup, Prod.Prompt.match.PC) %>%
  ungroup() %>%
  rename(PC = Prod.Prompt.match.PC) %>%
  mutate(Type = 'Real')

prompt.scramble <- data.scrambled %>%
  ungroup() %>%
  select(subj, month, VMSgroup, Prod.Prompt.match.PC) %>%
  rename(PC = Prod.Prompt.match.PC) %>%
  mutate(Type = 'Scrambled')

scramble.comparison.prompt <- rbind(prompt.real, prompt.scramble)

#shapiro.test(scramble.comparison.prompt$PC) # not normal

scramble.comparison.prompt %>% group_by(Type, VMSgroup) %>% summarise(meanscramble = mean(PC),
                                                                      sdscramble = sd(PC))

wilcox.test(subset(prompt.real, VMSgroup == "withVMS")$PC, 
            subset(prompt.scramble, VMSgroup == "withVMS")$PC, conf.int = T) # Significant, p<.001

cliff.delta(d = subset(prompt.real, VMSgroup == "withVMS")$PC, 
            f = subset(prompt.scramble, VMSgroup == "withVMS")$PC)   # .90 large

wilcox.test(subset(prompt.real, VMSgroup == "noVMS")$PC, 
            subset(prompt.scramble, VMSgroup == "noVMS")$PC, conf.int = T) # Significant, p<.05

cliff.delta(d = subset(prompt.real, VMSgroup == "noVMS")$PC, 
            f = subset(prompt.scramble, VMSgroup == "noVMS")$PC)   # .45 medium

```

```{r Figure 3a}

Figure3a <- 
  ggplot(scramble.comparison.prompt, 
         aes(x=Type, y=PC, shape = Type, color = VMSgroup))+
  geom_point(aes(group=VMSgroup), size = 5, position=position_dodge(-.2)) +
  stat_summary(fun.data=mean_cl_boot, geom = "pointrange", 
               aes(shape=Type, fill = VMSgroup), 
               size=1.5, position = position_dodge(-.2), colour = "grey32", show.legend = FALSE) +
  stat_summary(fun.y = mean, aes(group=VMSgroup, colour=VMSgroup), 
               geom='line', size=.8, position = position_dodge(-.2), show.legend = FALSE) +
  scale_shape_manual(name = "VMSgroup",
                      labels = c("noVMS", "withVMS"),
                       values=c(21,24)) +
  scale_colour_manual(name = "VMSgroup",
                     labels = c("noVMS", "withVMS"),
                      values = c("#F8766D","#00BFC4")) +
  xlab("Data type") +
  ylab('% Congruent CPs') +
  ggtitle("Caregiver Prompt") +
  theme_bw(base_size=15) +
  theme(axis.text = element_text(size = 20),
         axis.title = element_text(size=20),
         legend.title = element_blank(),
         legend.justification=c(1,1), legend.position=c(1,.99)) +
  guides(shape = FALSE,
         color = guide_legend(override.aes = list(shape=15)))
  
#plot(Figure3a)
```

A paired Wilcoxon Signed-Rank test showed that withVMS infants produced significantly more CPs to match caregiver prompts when the prompt was INREP (M=.71, SD=.24) than when it was OUTREP (M=.38, SD=.23; Est.Diff.=.35, r=.67, p=.001, 95% CI=[.15, .50]). 
```{r Input prompt inREP vs outREP comparison}

match.data.Prompt %>% 
  filter(VMSgroup == "withVMS") %>% 
  group_by(Type) %>% 
  summarise(meanmatch = mean(PC, na.rm = T),
            sdmatch = sd(PC, na.rm = T))

# shapiro.test(match.data.Prompt$PC) #normal

# t.test(subset(match.data.Prompt.spread, VMSgroup=="withVMS")$inREP,
# subset(match.data.Prompt.spread, VMSgroup=="withVMS")$outREP, conf.int=T, paired = T) # p<.001

wilcox.test(subset(match.data.Prompt.spread, VMSgroup=="withVMS")$inREP,
            subset(match.data.Prompt.spread, VMSgroup=="withVMS")$outREP, conf.int=T, paired = T) # p=.001, Est.Diff=.34

wilcoxonPairedR(x = subset(match.data.Prompt, VMSgroup == "withVMS")$PC,
                g = subset(match.data.Prompt, VMSgroup == "withVMS")$Type) # r= .67

```


```{r Figure 4a}

Figure4a <- 
  ggplot(match.data.Prompt, aes(x=Type, y=PC, shape = Type, color = VMSgroup))+
  geom_point(aes(group=VMSgroup), size = 5, position=position_dodge(-.2)) +
  geom_line(aes(group=subj), colour='grey63', position = position_nudge(x = -.05), linetype=6) +
  stat_summary(fun.data=mean_cl_boot, geom = "pointrange", 
               aes(shape=Type, fill = VMSgroup), 
               size=1.5, position = position_dodge(-.2), 
               colour = "grey32", show.legend = FALSE) +
  stat_summary(fun.y = mean, aes(group=VMSgroup, colour=VMSgroup), 
               geom='line', size=.8, position = position_dodge(-.2), show.legend = FALSE) +
  scale_shape_manual(values=c(23,25)) +
  xlab("Consonant Type") +
  ylab('% Congruent CPs') +
  ggtitle("Caregiver Prompt") +
  theme_bw(base_size=15) +
  theme(axis.text = element_text(size = 20),
        axis.title = element_text(size=20),
        legend.title = element_blank(),
        legend.justification=c(1,1), legend.position=c(1,.99)) +
  guides(shape = FALSE,
         color = guide_legend(override.aes = list(shape=15)))
  
#plot(Figure4a)
```

A Wilcoxon Rank Sum test revealed no difference between withVMS (M=.38, SD=.23) and noVMS infants’ CPs (M=.41, SD=.28) when the CP was not INREP (Est.Diff.=-.02, p=.741, δ=-.06, 95% CI=[-.2, .14]).  
```{r Input prompt outREP comparison}

match.data.Prompt %>% 
  filter(Type == "outREP") %>% 
  group_by(VMSgroup) %>% 
  summarise(meanMatch = mean(PC, na.rm=T),
            sdMatch = sd(PC, na.rm=T))

# shapiro.test(subset(match.data.Prompt, Type=='outREP')$PC)#normal

# t.test(PC ~ VMSgroup, subset(match.data.Prompt, Type=='outREP'), conf.int=T) # ns

wilcox.test(subset(match.data.Prompt, VMSgroup == "withVMS" & Type == "outREP")$PC, 
            subset(match.data.Prompt, VMSgroup == "noVMS" & Type == "outREP")$PC, conf.int = T) # ns

cliff.delta(d = subset(match.data.Prompt, VMSgroup == "withVMS" & Type == "outREP")$PC, 
            f = subset(match.data.Prompt, VMSgroup == "noVMS" & Type == "outREP")$PC)   # -.06 negligible

```


## Attended objects

On average, 56% of CPs (SD=.22) were produced while attending to an object, and 38% showed a phonological match with the attended object (SD=.27). To test whether the proportion of CPs that matched attended objects differed by VMS group, we ran a Wilcoxon Rank Sum test, comparing the proportion of matching CPs by VMS group. WithVMS infants had more matching CPs than noVMS infants (Est.Diff.=.28, δ=.46, p=.003, 95% CI=[.09, .43]; Table 4). 
```{r Object overview}

ProdData %>% summarize(meanObject = mean(nObj/CPtokens, na.rm=T), #.56
                   sdObject = sd(nObj/CPtokens, na.rm=T), #.22
                   Objectmatch = mean(Prod.Obj.match.PC, na.rm=T), # .39
                   SDobjectmatch = sd(Prod.Obj.match.PC, na.rm=T)) # 28

wilcox.test(subset(ProdData, VMSgroup == "withVMS")$Prod.Obj.match.PC, 
            subset(ProdData, VMSgroup == "noVMS")$Prod.Obj.match.PC, conf.int = T) # p<.01

cliff.delta(d = subset(ProdData, VMSgroup == "withVMS")$Prod.Obj.match.PC, 
            f = subset(ProdData, VMSgroup == "noVMS")$Prod.Obj.match.PC)   # .46 medium

```

We found that infants with a stable consonant repertoire matched their CPs to an attended object significantly more in the real dataset than in the scrambled one (.49(.27) vs. .24(.11), Est.Diff.=.28, δ=.61, p<.001, 95% CI=[.15, .40]; by Wilcoxon Rank Sum test), while infants who lacked such a repertoire did not (.25(.22) vs. .24(.16), Est.Diff.<.1, δ<.01, p=1, 95% CI=[-.17, .12]). 
```{r Object data vs scrambled data}

obj.real <- ProdData %>%
  select(subj, month, VMSgroup, Prod.Obj.match.PC) %>%
  ungroup() %>%
  rename(PC = Prod.Obj.match.PC) %>%
  mutate(Type = 'Real')

obj.scramble <- data.scrambled %>%
  ungroup() %>%
  select(subj, month, VMSgroup, Prod.Obj.match.PC) %>%
  rename(PC = Prod.Obj.match.PC) %>%
  mutate(Type = 'Scrambled')

scramble.comparison.object <- rbind(obj.real, obj.scramble)

scramble.comparison.object %>% group_by(Type, VMSgroup) %>% summarise(meanscramble = mean(PC, na.rm=T),
                                                                      sdscramble = sd(PC, na.rm=T))
# shapiro.test(scramble.comparison.object$PC) # not normal

wilcox.test(subset(obj.real, VMSgroup == "withVMS")$PC, 
            subset(obj.scramble, VMSgroup == "withVMS")$PC, conf.int = T) # p<.001

cliff.delta(d = subset(obj.real, VMSgroup == "withVMS")$PC, 
            f = subset(obj.scramble, VMSgroup == "withVMS")$PC)   # .61 large

wilcox.test(subset(obj.real, VMSgroup == "noVMS")$PC, 
            subset(obj.scramble, VMSgroup == "noVMS")$PC, conf.int = T) # ns

cliff.delta(d = subset(obj.real, VMSgroup == "noVMS")$PC, 
            f = subset(obj.scramble, VMSgroup == "noVMS")$PC)   # <.01 negligible

```

```{r Figure 3b}

Figure3b <- 
  ggplot(scramble.comparison.object, 
         aes(x=Type, y=PC, shape = Type, color = VMSgroup))+
  geom_point(aes(group=VMSgroup), size = 5, position=position_dodge(-.2)) +
  stat_summary(fun.data=mean_cl_boot, geom = "pointrange", 
               aes(shape=Type, fill = VMSgroup), 
               size=1.5, position = position_dodge(-.2), 
               colour = "grey32", show.legend = FALSE) +
  stat_summary(fun.y = mean, aes(group=VMSgroup, colour=VMSgroup), 
               geom='line', size=.8, position = position_dodge(-.2), show.legend = FALSE) +
  scale_shape_manual(name = "VMSgroup",
                     labels = c("noVMS", "withVMS"),
                     values=c(21,24)) +
  scale_colour_manual(name = "VMSgroup",
                      labels = c("noVMS", "withVMS"),
                      values = c("#F8766D","#00BFC4")) +
  xlab("Data type") +
  ylab('% Congruent CPs') +
  ggtitle("Attended Object") +
  theme_bw(base_size=15) +
  theme(axis.text = element_text(size = 20),
        axis.title = element_text(size=20),
        legend.title = element_blank(),
        legend.justification=c(1,1), legend.position=c(1,.99)) +
  guides(shape = FALSE,
         color = guide_legend(override.aes = list(shape=15)))
  
#plot(Figure3b)

multiplot(Figure3a, Figure3b, cols=2)
```


Using a paired Wilcoxon Signed Rank test, we compared the proportion of infants’ CPs that matched both the attended object and were part of their established VMS inventory (INREP; .57(.37)) with the proportion of CPs that matched the attended object but were not in the infant’s VMS inventory (OUTREP, .36(.32)). Similar to the caregiver prompts, infants produced more object-matching CPs when the object’s label provided a phonological match with their VMS inventory; this difference was marginally significant (Est.Diff.=.32, r=.42, p=.056, 95% CI=[-.01, .52]). 
```{r Object inREP vs outREP comparison}

# shapiro.test(subset(match.data.Object, VMSgroup=="withVMS")$PC) # not normal

match.data.Object %>% 
  filter(VMSgroup == "withVMS") %>% 
  group_by(Type) %>% 
  summarise(meanmatch = mean(PC, na.rm = T),
            sdmatch = sd(PC, na.rm = T))

wilcox.test(subset(match.data.Object.spread, VMSgroup=="withVMS")$inREP, 
            subset(match.data.Object.spread, VMSgroup=="withVMS")$outREP, conf.int=T, paired = T) # p=.056, Est Diff = .32

wilcoxonPairedR(x = subset(match.data.Object, VMSgroup == "withVMS")$PC,
                g = subset(match.data.Object, VMSgroup == "withVMS")$Type) # r= .42

```

```{r Figure 4b}

Figure4b <- ggplot(match.data.Object, aes(x=Type, y=PC)) +
  geom_point(aes(colour=VMSgroup, shape=Type), 
             size=5, position = position_dodge(-.2)) +
  geom_line(aes(group=subj), colour='grey63',position = position_nudge(x = -.05), linetype=6) +
  stat_summary(fun.data=mean_cl_boot, geom = "pointrange", 
               aes(shape=Type, fill= VMSgroup), colour = "gray32", size=1.5, 
               position = position_dodge(-.2), show.legend = FALSE) +
  stat_summary(fun.y = mean, aes(group=VMSgroup, colour=VMSgroup), 
               geom='line', size=.8, position = position_dodge(-.2), show.legend = FALSE) +
  scale_shape_manual(values=c(23,25)) +
  ylab("% Congruent CPs") +
  xlab("Consonant Type") +
  ggtitle("Attended Object") +
  theme_bw(base_size=15) +
   theme(axis.text.x = element_text(size = 20),
         axis.title.x = element_text(size=20),
         axis.title.y = element_text(size=20),
         axis.text.y = element_text(size=20),
         legend.title = element_blank(),
         legend.justification=c(1,1), legend.position=c(1,.99)) +
  guides(shape = FALSE,
         color = guide_legend(override.aes = list(shape=15)))

multiplot(Figure4a, Figure4b, cols=2)

#plot(Figure4)
```

Overall, 31% of OUTREP CPs matched an attended object (SD=.29), and this did not differ between VMS groups (Est.Diff.=-.08, δ=.13, p=.318, 95% CI=[-.07, .25]; withVMS: M=.36, SD=.32; noVMS: M=.25, SD=.22). 
```{r Object outREP comparison }

match.data.Object %>% filter(Type == "outREP") %>% summarise(meanMatch = mean(PC, na.rm=T),
                                                             sdMatch = sd(PC, na.rm=T))

wilcox.test(subset(match.data.Object, VMSgroup == "withVMS" & Type == "outREP")$PC, 
            subset(match.data.Object, VMSgroup == "noVMS" & Type == "outREP")$PC, conf.int = T) # p=.32, Est.Diff.=.08

cliff.delta(d = subset(match.data.Object, VMSgroup == "withVMS" & Type == "outREP")$PC, 
            f = subset(match.data.Object, VMSgroup == "noVMS" & Type == "outREP")$PC) # .13 negligible

match.data.Object %>% filter(Type == "outREP") %>% group_by(VMSgroup) %>% 
  summarise(meanMatch = mean(PC, na.rm=T),
            sdMatch = sd(PC, na.rm=T))
```


# Discussion


In a preliminary analysis of our home recordings, however, caregivers rarely responded to infants’ babble (<25%), leading us to forego such an analysis here. 
```{r CG responsiveness}

CPdata %>% filter(cgresponse == "None" | is.na(cgresponse)) %>% tally() #1404

1404/1916 #73%

```