---
title: "From babble to words: Infants’ early productions match words and objects in their environment"
author: "Catherine Laing & Elika Bergelson"
date: "19th January 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r Set up, include=FALSE}
library(ggplot2)
library(tidyverse)
library(readr)
library(forcats)
library(effsize)
library(rcompanion)
library(stringi)

multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
  library(grid)
  
  # Make a list from the ... arguments and plotlist
  plots <- c(list(...), plotlist)
  
  numPlots = length(plots)
  
  # If layout is NULL, then use 'cols' to determine layout
  if (is.null(layout)) {
    # Make the panel
    # ncol: Number of columns of plots
    # nrow: Number of rows needed, calculated from # of cols
    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
                     ncol = cols, nrow = ceiling(numPlots/cols))
  }
  
  if (numPlots==1) {
    print(plots[[1]])
    
  } else {
    # Set up the page
    grid.newpage()
    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))
    
    # Make each plot, in the correct location
    for (i in 1:numPlots) {
      # Get the i,j matrix positions of the regions that contain this subplot
      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))
      
      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
                                      layout.pos.col = matchidx$col))
    }
  }
}
```

```{r data, include=FALSE}
source("ProductionData_randsubj.R")
source("Data_scrambling_randsubj.R")
```

# Methods

Thirteen infants were withVMS according to these criteria. Of the remaining 29 infants, fourteen had low production at 0;10 and were classed as noVMS (M=8.5 overall tokens, SD=8.46, R=0-28). The remaining 17 infants showed signs of a developing VMS consonant at 0;10: a given consonant was dominant in the session (defined here as ≥20 tokens of the same consonant accounting for >20% of overall consonant production, cf. Athari, Wang, Dey & Rvachew, 2019), but did not reach our initial criteria of  ≥50 tokens (mean n=29.5, SD = 8.64, R=20-48; Mean %=.50, SD=.17, R=.23-.9). As this suggests that these infants were on the cusp of establishing a VMS, we re-tested this subset at 0;11 to ensure our VMS groups were sufficiently differentiated. Infants who showed low/high production at 0;11 were then grouped as noVMS (n=6) or withVMS (n=8), respectively, as per our initial criteria. Finally, three infants were consistent in producing the same consonant stably at 0;10 and 0;11 (≥20 productions of a consonant type, accounting for >20% of all consonant tokens at both 0;10 and 0;11), thereby demonstrating ‘well- practiced and longitudinally stable vocal productions’ (McCune & Vihman, 2001, p.671), and were also categorized as withVMS. 

```{r Establishing VMS groups}

vmscount %>% filter(month_chron == 10) %>% 
  mutate(criterion1 = ifelse(n > 49, T, F)) %>% 
  filter(criterion1 == T) %>% group_by(subj) %>% 
  tally() %>% 
  ungroup() %>% 
  tally()   # 13 infants


vmscount %>% filter(month_chron == 10) %>% 
  mutate(criterion2 = ifelse(n < 20, T, F)) %>% 
  filter(criterion2 == T) %>% group_by(subj) %>% 
  tally() %>% 
  filter(n == 5) %>%
  tally()   # 14 infants had no stable consonants 

vmscount %>% filter(subj %in% vmssubset1$subj & VMSgroup == "noVMS") %>%
  ungroup() %>%
  summarise(meantotal = mean(totalprod), # 8.5
            sdtotal = sd(totalprod),     # 8.46
            mintotal = min(totalprod),   # 0
            maxtotal = max(totalprod))   # 28

vmscount %>% filter(month_chron == 10) %>%
  mutate(criterion3 = ifelse(n > 19 & n < 50 & subj %in% vms_subset2$subj, T, F)) %>%
  filter(criterion3 == T) %>%
  group_by(subj) %>%
  mutate(PC = n/totalprod) %>%
  ungroup() %>%
  summarise(meanN = mean(n),  # 29.5
         sdN = sd(n),         # 8.64
         minN = min(n),       # 20
         maxN = max(n),       # 48
         meanPC = mean(PC),   # .5
         sdPC = sd(PC),       # .17
         minPC = min(PC),     # .23
         maxPC = max(PC))     # .9

vmscount %>% 
  group_by(subj, VMSgroup, month_chron, n) %>% 
  filter(subj %in% vms_subset2$subj & 
           month_chron == 11 & 
           n < 20) %>%
  ungroup() %>%
  group_by(subj) %>%
  tally() %>%
  filter(n == 5) # 6 infants
  
vmscount %>% 
  group_by(subj, VMSgroup, month_chron, n) %>% 
  filter(subj %in% vms_subset2$subj & 
           month_chron == 11 & 
           n > 50) %>%
  ungroup() %>%
  group_by(subj) %>%
  tally() # 8 infants

vmscount %>% 
  group_by(subj, VMSgroup, month_chron, n) %>% 
  filter(subj %in% vms_subset2$subj &
           month_chron == 11 & 
           n < 50 & n >19 &
           inREP == T) %>%
  #view()
  ungroup() %>%
  group_by(subj) %>%
  tally()   # 3 infants

```

```{r Table 1}

vmstotal %>% group_by(sex) %>% filter(month == 10 & VMSgroup == "withVMS") %>% tally() # 13(5)
vmstotal %>% group_by(sex) %>% filter(month == 11 & VMSgroup == "withVMS") %>% tally() # 11(8)
vmstotal %>% group_by(sex) %>% filter(month == 10 & VMSgroup == "noVMS") %>% tally() # 14(5)
vmstotal %>% group_by(sex) %>% filter(month == 11 & VMSgroup == "noVMS") %>% tally() # 6(3)
vmstotal %>% group_by(sex) %>% filter(month == 11) %>% tally() # 17(11)

```



Descriptively, 24/44 infants had at least one VMS consonant; 9 had two VMS in repertoire (5 females) two infants (one female) had three. The most common VMS consonant was /p,b/, which featured in 16 infants’ repertoires. Ten infants had /t,d/ in repertoire, and eight had /k, ɡ/ (see S1 and ST1). 

```{r Summarizing VMS consonants & ST1}

vmstotal %>% filter(VMSgroup == "withVMS") %>% group_by(sex) %>% tally() #24
vmscount %>% filter(Group == "Multi") %>% group_by(subj, VMS, sex) %>% tally() #11
vmscount %>% filter(Group == "One") %>% group_by(subj, VMS, sex) %>% tally() #13


vmstotal %>% filter(VMSgroup == "noVMS") %>% group_by(sex) %>% tally() #20 (8)
vmstotal %>% filter(VMS == "p b") %>% group_by(sex) %>% tally() # 8 (6)
vmstotal %>% filter(VMS == "t d") %>% group_by(sex) %>% tally() # 4 (1)
vmstotal %>% filter(VMS == "k g") %>% group_by(sex) %>% tally() # 1 (0)
vmstotal %>% filter(VMS == "p b m") %>% group_by(sex) %>% tally() # 1 (0)
vmstotal %>% filter(VMS == "p b t d") %>% group_by(sex) %>% tally() # 1 (1)
vmstotal %>% filter(VMS == "p b k g") %>% group_by(sex) %>% tally() # 4(2)
vmstotal %>% filter(VMS == "t d k g") %>% group_by(sex) %>% tally() # 2(1)
vmstotal %>% filter(VMS == "t d n") %>% group_by(sex) %>% tally() # 1(1)
vmstotal %>% filter(VMS == "p b t d n") %>% group_by(sex) %>% tally() # 1(0)
vmstotal %>% filter(VMS == "p b t d k g") %>% group_by(sex) %>% tally() # 1 (1)

```


Every consonant that the infant produced in their hour-long video at month 10 (n=27) or 11 (n=17)

```{r Infant age in months; Table 3}

vmstotal %>% group_by(month) %>% tally()

vmstotal %>% group_by(VMSgroup) %>% filter(month == 10) %>% tally()

vmstotal %>% group_by(VMSgroup) %>% filter(month == 11) %>% tally()

```


Average duration between caregiver prompt and infant CP was 6.07s (SD=3.8); results were consistent when we reanalyzed the data with only prompts from <10s before the infants’ CP. 

```{r Determining timestamp variability for CG prompt data}

CPdata %>% 
  filter(timediff <= 15000) %>%
  summarise(meanprompt = mean(timediff),
            medianprompt = median(timediff),
            maxprompt = max(timediff),
            minprompt = min(timediff),
            sdprompt = sd(timediff))

```

51% (n=1022) of CPs were preceded by audible caregiver speech in this timeframe. The majority of these (74%) included a noun.

```{r Counting data for CGprompt 1}

CPdata %>% tally() # 1916 CPs

CPdata %>% filter(cgprompt != "[none]" & 
                    cgprompt != "[unclear]" & 
                    cgprompt != "[unavailable]" & 
                    cgprompt != "[imitation]" &
                    cgprompt != "[media]") %>% 
  tally() # 951

CPdata %>%
  group_by(subj) %>%
  filter(cgprompt != "[none]" & 
                    cgprompt != "[unclear]" & 
                    cgprompt != "[unavailable]" & 
                    cgprompt != "[imitation]" &
                    cgprompt != "[media]") %>% 
  tally() %>% # 951
  summarise(meanprompt = mean(n),
            sdprompt = sd(n))

951/1916 # 51%

CPdata %>% filter(wordclass != "[none]") %>% tally() # 951
CPdata %>% filter(wordclass == "N") %>% tally() # 708

708/951  # 74%

951-708

```



Attended objects accounted for 62% of our CP data (n=1,179 CPs). After removing instances when the infant was attending to an object but its identity or label was unclear (n=78), we were left with 1,101 data points for the attended object analysis. 

```{r Counting data for AttObj}

CPdata %>% filter(object != "[none]" & object != "[unclear]" & object != "[out of view]") %>% tally() # 1179

CPdata %>% filter(object != "[none]" & object != "[unclear]" & object != "[out of view]") %>% 
  group_by(subj) %>% 
  tally() %>%
  summarise(meanObj = mean(n),
            sdObj = sd(n))


1179/1916 # 62%

CPdata %>% filter(object == "[toy]") %>% tally() # 78

1179-78

```

For both caregiver input and attended object, in around 5% of cases (40 caregiver prompts and 58 attended objects) this meant there was no codable consonant for a word (e.g. neither wow nor  whisk have a supra-glottal consonant with full closure in syllable-initial position).
```{r Counting data for CGprompt 2}

CPdata %>% filter(!grepl("^\\[",cgprompt) & C1Prompt == "none") %>% tally()  # 40 instances of CGprompts with no supraglottal consonant

CPdata %>% filter(!grepl("^\\[", object) & C1Obj == "none") %>% tally()  # 58 instances of Objects with no supraglottal consonant


 1916 - (40+58) # 1818
 (40+58)/1916 # 5%
```

# Results

An initial analysis of infant age, sex, and maternal education showed that these variables did not account for significant variance in the number of CP types or tokens, in audio-recordings or in videos, and thus data are collapsed across these dimensions in all further analysis. See S3 for further details.

```{r Data Analysis Plan - confounding variables - Audio data}

# CP Tokens 

#shapiro.test(vmstotal$audiotokens) # Not normal; 
                                    # log(audiotokens+1) also not normal so stick with non-transformed data and Non-para tests

wilcox.test(audiotokens ~ month, vmstotal) # ns p=.75

wilcox.test(audiotokens ~ sex, vmstotal) # ns p=.5

cor.test(vmstotal$MOTedulevel, vmstotal$audiotokens, method = 'spearman') # ns p=.3

# CP types

#shapiro.test(CPtypes_audio$CPtypes) # not normal

wilcox.test(CPtypes ~ sex, CPtypes_audio) # ns, p=.44
wilcox.test(CPtypes ~ month, CPtypes_audio) #ns, p=.61

cor.test(CPtypes_audio$MOTedulevel, CPtypes_audio$CPtypes, method = 'spearman') # ns p=.39  

```


```{r Data Analysis Plan - counfounding variables - Video data}

shapiro.test(ProdData$CPtokens) # Not normal
shapiro.test(ProdData$CPtypes) # Not normal

wilcox.test(CPtokens ~ sex, ProdData) # p=.66

wilcox.test(CPtypes ~ sex, ProdData) # p=.97

wilcox.test(CPtokens ~ month, ProdData) # p=.596

wilcox.test(CPtypes ~ month, ProdData) # p=.36

cor.test(ProdData$MOTedulevel, ProdData$CPtokens, method = 'spearman') # p=.193
cor.test(ProdData$MOTedulevel, ProdData$CPtypes, method = 'spearman') # p=.886
```

None of the infants were outliers in our dataset (defined as producing more/less than 3 standard deviations from the mean CP types/tokens).
```{r outliers}

ProdData %>% mutate(outlier = (CPtokens > 
                      (mean(CPtokens, na.rm=T) + 
                         3*(sd(CPtokens, na.rm=T))) |
                      CPtokens <
                      (mean(CPtokens, na.rm=T) - 
                         3*(sd(CPtokens, na.rm=T))))) %>%
  filter(outlier == TRUE)  # no outliers

ProdData %>% mutate(outlier = (CPtypes > 
                      (mean(CPtypes, na.rm=T) + 
                         3*(sd(CPtypes, na.rm=T))) |
                      CPtypes <
                      (mean(CPtypes, na.rm=T) - 
                         3*(sd(CPtypes, na.rm=T))))) %>%
  filter(outlier == TRUE)  # no outliers

```

# Infant production: Audio recordings

On average, infants produced 96 consonant tokens in the half-hour segment of the daylong audio-recordings we analyzed. Variability was high (Range=0=610; see Table 4); three infants (2 females) produced zero consonants. The most prominent consonant was /p,b/ (n=1716), followed by /t,d/ (n=1364) and /k, g/ (n=745). 
```{r Audio recordings summary}

vmscount %>% 
  filter(month_chron == month) %>%
  group_by(subj) %>% 
  summarise(total = sum(n)) %>% 
  ungroup() %>% 
  summarise(meanprod = mean(total), # 95.8
            sdprod = sd(total),  # 117.12
            minprod = min(total), # 0
            maxprod = max(total)) # 610

vmscount %>% group_by(subj) %>% summarise(total = sum(n)) %>% filter(total == 0) %>% tally() #3 infant with 0 productions

vmscount %>% 
  filter(month_chron == month) %>% 
  group_by(Consonant) %>% summarise(total = sum(n))

```

```{r Table 4 Audio}

# Audio

vmstotal %>% summarise(meantyp = mean(audiotypes),
                       sdtyp = sd(audiotypes))

vmstotal %>% group_by(VMSgroup) %>%
  summarise(meantyp = mean(audiotypes),
                       sdtyp = sd(audiotypes))

vmscount %>% filter(VMSgroup == "withVMS") %>% 
  group_by(subj) %>% 
  summarise(total = sum(n)) %>% 
  ungroup() %>%
  summarise(meanprod = mean(total),
                          sdprod = sd(total))

vmscount %>% filter(VMSgroup == "noVMS") %>% 
  group_by(subj) %>% 
  summarise(total = sum(n)) %>% 
  ungroup() %>%
  summarise(meanprod = mean(total),
                          sdprod = sd(total))
```

```{r Table 4 Video}

# Video

CPdata %>% group_by(subj) %>% 
  summarise(videotypes = n_distinct(ctype)) %>% ungroup() %>%
  summarise(Mvideotypes = mean(videotypes),
            SDvideotypes = sd(videotypes))

CPdata %>% group_by(subj) %>% 
  filter(VMSgroup == "withVMS") %>%
  summarise(videotypes = n_distinct(ctype)) %>% ungroup() %>%
  summarise(Mvideotypes = mean(videotypes),
            SDvideotypes = sd(videotypes))

CPdata %>% group_by(subj) %>% 
  filter(VMSgroup == "noVMS") %>%
  summarise(videotypes = n_distinct(ctype)) %>% ungroup() %>%
  summarise(Mvideotypes = mean(videotypes),
            SDvideotypes = sd(videotypes))

CPdata %>% group_by(subj) %>% tally() %>%
  ungroup() %>%
  summarise(Mvideotok = mean(n),
            SDvideotok = sd(n))

CPdata %>% filter(VMSgroup == "withVMS") %>%
  group_by(subj) %>% tally() %>%
  ungroup() %>%
  summarise(Mvideotok = mean(n),
            SDvideotok = sd(n))

CPdata %>% filter(VMSgroup == "noVMS") %>%
  group_by(subj) %>% tally() %>%
  ungroup() %>%
  summarise(Mvideotok = mean(n),
            SDvideotok = sd(n))


# Input prompt match

table(is.na(ProdData$Prod.Prompt.match.PC)) # all 44
ProdData %>% summarize(Promptmatch = mean(Prod.Prompt.match.PC, na.rm=T),
                   SDPromptmatch = sd(Prod.Prompt.match.PC, na.rm=T))

ProdData %>% group_by(VMSgroup) %>% summarize(Promptmatch = mean(Prod.Prompt.match.PC, na.rm=T),
                   SDPromptmatch = sd(Prod.Prompt.match.PC, na.rm=T))

# Attended object match
table(is.na(ProdData$Prod.Obj.match.PC)) # only 43 infants
ProdData %>% summarize(Objmatch = mean(Prod.Obj.match.PC, na.rm=T),
                   SDObjmatch = sd(Prod.Obj.match.PC, na.rm=T))

ProdData %>% group_by(VMSgroup) %>% summarize(Objmatch = mean(Prod.Obj.match.PC, na.rm=T),
                   SDObjmatch = sd(Prod.Obj.match.PC, na.rm=T))

```

As expected, withVMS infants produced more babble in the audio recordings, and this difference was significant (Est.Diff=-130, p<.001 by Wilcoxon Signed-Rank test)

```{r comparing production across groups in audio recordings}

shapiro.test(vmstotal$audiotypes) # not normal

wilcox.test(subset(vmstotal, VMSgroup == "withVMS")$audiotypes, subset(vmstotal, VMSgroup == "noVMS")$audiotypes, conf.int=T) # p<.001

cliff.delta(d = subset(vmstotal, VMSgroup == "withVMS")$audiotypes, 
     f = subset(vmstotal, VMSgroup == "noVMS")$audiotypes)   # .69 large

shapiro.test(vmstotal$audiotokens) # not normal

wilcox.test(subset(vmstotal, VMSgroup == "withVMS")$audiotokens, subset(vmstotal, VMSgroup == "noVMS")$audiotokens, conf.int=T) # p<.001

cliff.delta(d = subset(vmstotal, VMSgroup == "withVMS")$audiotokens, 
     f = subset(vmstotal, VMSgroup == "noVMS")$audiotokens)   # .99 large



```

```{r Figure 1}

groups <- c("withVMS" = "with-VMS",
            "noVMS" = "no-VMS")

Figure1a <- ggplot(subset(vmscount, VMSgroup == "withVMS" & month_chron == month), aes(x=Consonant, y=log10(n))) +
  stat_summary(fun.y=mean, geom = "point", aes(group = subj, colour = Consonant), shape=1, size=2, 
               position = position_jitter(width = .1, height=0)) +
  stat_summary(fun.data=mean_cl_boot, geom = "pointrange", shape=2, size=1) +
  scale_x_discrete(limits = c("p", "t", "k", "m", "n"),
                   labels = c("pb", "td", "kg", "m", "n")) +
  theme_bw(base_size=15) +
  scale_y_continuous(breaks=c(0, .301, 1, 1.699, 2.69897), labels=c(0, 2, 10, 50, 500)) +
  xlab("Consonant Type") +
  ylab("n") +
  theme(legend.position = "none",
        axis.title = element_text(size=22),
           axis.text = element_text(size=22),
        strip.text = element_text(face="bold", size=22,lineheight=7.0)) +
  facet_wrap(~VMSgroup, labeller = as_labeller(groups))

Figure1b <- ggplot(subset(vmscount, VMSgroup == "noVMS" & month_chron == month), aes(x=Consonant, y=log10(n))) +
  stat_summary(fun.y=mean, geom = "point", aes(group = subj, colour = Consonant), shape=1, size=2,
               position = position_jitter(width = .1, height=0)) +
  stat_summary(fun.data=mean_cl_boot, geom = "pointrange", shape=2, size=1) +
  scale_x_discrete(limits = c("p", "t", "k", "m", "n"),
                   labels = c("pb", "td", "kg", "m", "n")) +
  theme_bw(base_size=15) +
  scale_y_continuous(breaks=c(0, .301, 1, 1.477, 1.699, 2), labels=c(0, 2, 10, 30, 50, 100)) +
  xlab("Consonant Type") +
  ylab("n") +
  theme(legend.position = "none",
        axis.title = element_text(size=22),
           axis.text = element_text(size=22),
        strip.text = element_text(face="bold", size=22,lineheight=7.0)) +
  facet_wrap(~VMSgroup, labeller = as_labeller(groups))

multiplot(Figure1a, Figure1b, cols=2)
```
  
# Infant Production: Video recordings


On average, infants produced 44 CP tokens in the hour-long video recordings and 4.34 CP types. Once again, withVMS infants produced significantly more tokens than noVMS infants, while the number of CP types was marginally higher for withVMS infants.

```{r Effect of VMS group on CP types/tokens}

wilcox.test(subset(ProdData, VMSgroup == "withVMS")$CPtokens, subset(ProdData, VMSgroup == "noVMS")$CPtokens, conf.int=T) # p<.001

cliff.delta(d = subset(ProdData, VMSgroup == "withVMS")$CPtokens, 
     f = subset(ProdData, VMSgroup == "noVMS")$CPtokens)   # .68 large

ProdData %>% group_by(VMSgroup) %>% summarise(meantok = mean(CPtokens))

wilcox.test(subset(ProdData, VMSgroup == "withVMS")$CPtypes, subset(ProdData, VMSgroup == "noVMS")$CPtypes, conf.int=T) # p=.06

cliff.delta(d = subset(ProdData, VMSgroup == "withVMS")$CPtypes, 
     f = subset(ProdData, VMSgroup == "noVMS")$CPtypes)   # .32 small

ProdData %>% group_by(VMSgroup) %>% summarise(meantyp = mean(CPtypes))

```

As expected, withVMS infants produced their INREP consonants significantly more often (mean tokens=23.3, SD=21.7) than their OUTREP consonants (mean tokens=8.7, SD= 7.5; Est.Diff.=-9.6, p<.01). 
  
```{r calcuating number of in vs out VMS consonants in video recordings}

compareRecordings %>% filter(VMS.Prod.Cong == T) %>% summarise(meanOverall = mean(meanProd),
                                                               sdOverall = sd(meanProd))

compareRecordings %>% filter(VMS.Prod.Cong == F) %>% summarise(meanOverall = mean(meanProd),
                                                               sdOverall = sd(meanProd))

shapiro.test(compareRecordings$meanProd) # not normal

wilcox.test(compareRecordings.spread$VMS.Prod.Cong, compareRecordings.spread$VMS.Prod.Incong, paired=T, conf.int=T)

wilcoxonPairedR(x = compareRecordings$meanProd,
                g = compareRecordings$VMS.Prod.Cong) # r= .63

```


```{r Figure 2}
Figure2a <- ggplot(ProdData, aes(x=VMSgroup, y=CPtokens)) +
  geom_violin(aes(fill = VMSgroup), alpha=.3) +
  stat_summary(fun.y=mean, geom = "point", aes(group = subj), shape=1, size=2.5, position=position_jitter(height = 0, width = 0.01)) +
  stat_summary(fun.data=mean_cl_boot, geom = "pointrange", colour='red', shape=17, size=1) +
  scale_x_discrete(limits = c("withVMS", "noVMS"), labels = c("with-VMS", "no-VMS")) +
    theme_bw(base_size=15) +
     theme(axis.text = element_text(size=18),
           axis.title = element_text(size=18),
           legend.position="none") +
  ylab("n CP tokens") +
  xlab('')
Figure2b <- ggplot(ProdData, aes(x=VMSgroup, y=CPtypes)) +
    geom_violin(aes(fill = VMSgroup), alpha=.3) +
  stat_summary(fun.y=mean, geom = "point", aes(group = subj), shape=1, size=2.5, position=position_jitter(height = 0, width = .1)) +
  stat_summary(fun.data=mean_cl_boot, geom = "pointrange", colour='red', shape=17, size=1) +
  scale_x_discrete(limits = c("withVMS", "noVMS"), labels = c("with-VMS", "no-VMS")) +
  coord_cartesian(ylim=c(0, 8)) +
  scale_y_continuous(breaks=seq(0, 8, 2))  +
   ylab("n CP types") +
  xlab('') +
  theme_bw(base_size=15) +
  theme(legend.position="none",
        axis.title = element_text(size=18),
        axis.text = element_text(size=18))


multiplot(Figure2a, Figure2b, cols=2)
```


## Input prompts

Overall, 1441 CPs in the data were produced alongside either an input prompt or an attended object; 512 CPs were produced alongside both an input prompt and an attended object.  
```{r Overview of context and CPs}

CPdata %>% filter(C1Prompt != "none") %>% tally() # 911

CPdata %>% mutate(nostim = ifelse(C1Obj == "none" & C1Prompt == "none", T, F)) %>%
  filter(nostim == FALSE) %>% tally() # 1441

CPdata %>% mutate(bothstim = ifelse(C1Obj != "none" & C1Prompt != "none", T, F)) %>%
  filter(bothstim == TRUE) %>% tally() # 512
```




On average, 50% of the infants' CPs were preceded by caregiver speech (SD=.18), and 49% of these matched the salient word in this preceding segment (SD=.23). ). We ran a Wilcoxon Rank Sum test Tto determine whether the proportion of each infant’s congruent CPs differed by VMS group, we ran a one-way ANOVA.. This difference was  not significant VMS group accounted for a significant amount of variance in our data (Est.Diff.=-.11, p=.11F(1,42)=4.27, p<.05; see Tablele 4);  withVMS infants’ CPs matched input promptcaregiver prompts 145%  more often than noVMS infants’ did (.55(.17) vs. .41(.28)). 
```{r Input prompt overview}

ProdData %>% summarize(meanPrompt = mean(nPrompt/CPtokens, na.rm=T), #.50
                   sdPrompt = sd(nPrompt/CPtokens, na.rm=T), #.18
                   Promptmatch = mean(Prod.Prompt.match.PC, na.rm=T), #.49
                   SDPromptmatch = sd(Prod.Prompt.match.PC, na.rm=T)) #.23

#shapiro.test(ProdData$Prod.Prompt.match.PC) # p=.048

#t.test(Prod.Prompt.match.PC ~ VMSgroup, ProdData) # p=.057

wilcox.test(subset(ProdData, VMSgroup=="withVMS")$Prod.Prompt.match.PC, 
            subset(ProdData, VMSgroup=="noVMS")$Prod.Prompt.match.PC, conf.int=T) # p=.11, EstDiff=-.12

cliff.delta(d = subset(ProdData, VMSgroup=="withVMS")$Prod.Prompt.match.PC, 
     f = subset(ProdData, VMSgroup=="noVMS")$Prod.Prompt.match.PC)   # .28 small


ProdData %>% group_by(VMSgroup) %>% summarise(meanInput = mean(Prod.Prompt.match.PC),
                                              sdInput = sd(Prod.Prompt.match.PC))

```

WithVMS infants’ CPs matched the consonantal properties of their caregivers’ speech significantly more in the real data set than in the scrambled one (.55(.17) vs. .22(.14), Est.Diff.=.33, p<.001 by Wilcoxon Test), as did noVMS infants’ CPs (.41(.28) vs. .19(.15), Est.Diff.=.24, p<.01). 
```{r Input prompt data vs scrambled data}

prompt.real <- ProdData %>%
  select(subj, month, VMSgroup, Prod.Prompt.match.PC) %>%
  ungroup() %>%
  rename(PC = Prod.Prompt.match.PC) %>%
  mutate(Type = 'Real')

prompt.scramble <- data.scrambled %>%
  ungroup() %>%
  select(subj, month, VMSgroup, Prod.Prompt.match.PC) %>%
  rename(PC = Prod.Prompt.match.PC) %>%
  mutate(Type = 'Scrambled')

scramble.comparison.prompt <- rbind(prompt.real, prompt.scramble)

#shapiro.test(scramble.comparison.prompt$PC) # not normal

scramble.comparison.prompt %>% group_by(Type, VMSgroup) %>% summarise(meanscramble = mean(PC),
                                                                      sdscramble = sd(PC))

# Now compare groups

wilcox.test(subset(prompt.real, VMSgroup == "withVMS")$PC, 
            subset(prompt.scramble, VMSgroup == "withVMS")$PC, conf.int = T) # Significant, p<.001

cliff.delta(d = subset(prompt.real, VMSgroup == "withVMS")$PC, 
     f = subset(prompt.scramble, VMSgroup == "withVMS")$PC)   # .90 large

wilcox.test(subset(prompt.real, VMSgroup == "noVMS")$PC, 
            subset(prompt.scramble, VMSgroup == "noVMS")$PC, conf.int = T) # Significant, p<.05

cliff.delta(d = subset(prompt.real, VMSgroup == "noVMS")$PC, 
     f = subset(prompt.scramble, VMSgroup == "noVMS")$PC)   # .45 medium

```

```{r Figure 3a}

Figure3a <- 
  ggplot(scramble.comparison.prompt, aes(x=Type, y=PC, shape = Type, color = VMSgroup))+
  geom_point(aes(group=VMSgroup), size = 5, position=position_dodge(-.2)) +
  stat_summary(fun.data=mean_cl_boot, geom = "pointrange", aes(shape=Type, fill = VMSgroup), 
               size=1.5, position = position_dodge(-.2), colour = "grey32", show.legend = FALSE) +
  stat_summary(fun.y = mean, aes(group=VMSgroup, colour=VMSgroup), geom='line', size=.8, position = position_dodge(-.2), show.legend = FALSE) +
    scale_shape_manual(name = "VMSgroup",
                      labels = c("no-VMS", "with-VMS"),
                       values=c(21,24)) +
    scale_colour_manual(name = "VMSgroup",
                     labels = c("no-VMS", "with-VMS"),
                      values = c("#F8766D","#00BFC4")) +
  xlab("Data type") +
  ylab('% Congruent CPs') +
  ggtitle("Caregiver Prompt") +
  theme_bw(base_size=15) +
   theme(axis.text = element_text(size = 20),
         axis.title = element_text(size=20),
         legend.title = element_blank(),
         legend.justification=c(1,1), legend.position=c(1,.99)) +
  guides(shape = FALSE,
         color = guide_legend(override.aes = list(shape=15)))
  
#plot(Figure3a)
```

A paired Wilcoxon Signed-Rank test showed that withVMS infants produced significantly more CPs to match caregiver prompts when the prompt was INREP (M=.71, SD=.25) than when it was OUTREP (M=.38, SD=.23; Est.Diff.=.35, p=.001).  
```{r Input prompt inREP vs outREP comparison}

#since value of PC for noVMS is all NA, no need to even subset
shapiro.test(subset(match.data.Prompt, VMSgroup=="withVMS")$PC) #normal

# t.test(subset(match.data.Prompt.spread, VMSgroup=="withVMS")$inREP,
# subset(match.data.Prompt.spread, VMSgroup=="withVMS")$outREP, conf.int=T, paired = T) # p<.001

wilcox.test(subset(match.data.Prompt.spread, VMSgroup=="withVMS")$inREP,
subset(match.data.Prompt.spread, VMSgroup=="withVMS")$outREP, conf.int=T, paired = T) # p=.001, Est.Diff=.34

wilcoxonPairedR(x = subset(match.data.Prompt, VMSgroup == "withVMS")$PC,
                g = subset(match.data.Prompt, VMSgroup == "withVMS")$Type) # r= .67

match.data.Prompt %>% 
  filter(VMSgroup == "withVMS") %>% 
  group_by(Type) %>% 
  summarise(meanmatch = mean(PC, na.rm = T),
            sdmatch = sd(PC, na.rm = T))

```


```{r Figure 4a}
Figure4a <- 
  ggplot(match.data.Prompt, aes(x=Type, y=PC, shape = Type, color = VMSgroup))+
  geom_point(aes(group=VMSgroup), size = 5, position=position_dodge(-.2)) +
  geom_line(aes(group=subj), colour='grey63', position = position_nudge(x = -.05), linetype=6) +
  stat_summary(fun.data=mean_cl_boot, geom = "pointrange", aes(shape=Type, fill = VMSgroup), 
               size=1.5, position = position_dodge(-.2), colour = "grey32", show.legend = FALSE) +
  stat_summary(fun.y = mean, aes(group=VMSgroup, colour=VMSgroup), geom='line', size=.8, position = position_dodge(-.2), show.legend = FALSE) +
    scale_shape_manual(values=c(23,25)) +
  xlab("Consonant Type") +
  ylab('% Congruent CPs') +
  ggtitle("Caregiver Prompt") +
  scale_colour_discrete(breaks=c("noVMS", "withVMS"),
                      labels = c("no-VMS", "with-VMS")) +
  theme_bw(base_size=15) +
   theme(axis.text = element_text(size = 20),
         axis.title = element_text(size=20),
         legend.title = element_blank(),
         legend.justification=c(1,1), legend.position=c(1,.99)) +
  guides(shape = FALSE,
         color = guide_legend(override.aes = list(shape=15)))
  
#plot(Figure4a)
```

A Wilcoxon Rank Sum test revealed no difference between withVMS (M=.38, SD=.23) and noVMS infants’ CPs (M=.41, SD=.28) when the CP was not INREP (Est.Diff.=.02, p=.74).   
```{r Input prompt outREP comparison}

match.data.Prompt %>% 
  filter(Type == "outREP") %>% 
  summarise(meanMatch = mean(PC, na.rm=T),
              sdMatch = sd(PC, na.rm=T))

match.data.Prompt %>% 
  filter(Type == "outREP") %>% 
  group_by(VMSgroup) %>% 
  summarise(meanMatch = mean(PC, na.rm=T),
            sdMatch = sd(PC, na.rm=T))

shapiro.test(subset(match.data.Prompt, Type=='outREP')$PC)#normal

#t.test(PC ~ VMSgroup, subset(match.data.Prompt, Type=='outREP'), conf.int=T) # ns

wilcox.test(subset(match.data.Prompt, VMSgroup == "withVMS" & Type == "outREP")$PC, 
            subset(match.data.Prompt, VMSgroup == "noVMS" & Type == "outREP")$PC, conf.int = T) # ns

cliff.delta(d = subset(match.data.Prompt, VMSgroup == "withVMS" & Type == "outREP")$PC, 
     f = subset(match.data.Prompt, VMSgroup == "noVMS" & Type == "outREP")$PC)   # -.06 negligible

```


## Attended objects

On average, 56% of CPs (SD=.22) were produced while attending to an object, and 38% showed a phonological match with the attended object (SD=.27). To test whether the proportion of CPs that matched attended objects differed by VMS group, we ran a Wilcoxon Rank Sum test with proportion of matching CPs as our dependent variable and VMS group as a factor. This showed that VMS group accounted for a significant amount of variance in the data (Est.Diff.=-.28, p<.01; Table 5). 
```{r Object overview}

CPdata %>% filter(C1Obj != "none") %>% tally() # 1042
ProdData %>% summarize(meanObject = mean(nObj/CPtokens, na.rm=T), #.56
                   sdObject = sd(nObj/CPtokens, na.rm=T), #.22
                   Objectmatch = mean(Prod.Obj.match.PC, na.rm=T), # .38
                   SDobjectmatch = sd(Prod.Obj.match.PC, na.rm=T)) # 27


wilcox.test(subset(ProdData, VMSgroup == "withVMS")$Prod.Obj.match.PC, 
            subset(ProdData, VMSgroup == "noVMS")$Prod.Obj.match.PC, conf.int = T) # p<.01

cliff.delta(d = subset(ProdData, VMSgroup == "withVMS")$Prod.Obj.match.PC, 
     f = subset(ProdData, VMSgroup == "noVMS")$Prod.Obj.match.PC)   # .46 medium

```

We found that infants with a stable consonant repertoire matched their CPs to an attended object significantly more in the real dataset than in the scrambled one (.49(.27) vs. .24(.11), Est.Diff.=.28, p<.001; by Wilcoxon Signed-Rank test), while infants who lacked such a repertoire did not (.25(.22) vs. .25(.16), Est.Diff.<.1, p=1). 
```{r Object data vs scrambled data}

obj.real <- ProdData %>%
  select(subj, month, VMSgroup, Prod.Obj.match.PC) %>%
  ungroup() %>%
  rename(PC = Prod.Obj.match.PC) %>%
  mutate(Type = 'Real')

obj.scramble <- data.scrambled %>%
  ungroup() %>%
  select(subj, month, VMSgroup, Prod.Obj.match.PC) %>%
  rename(PC = Prod.Obj.match.PC) %>%
  mutate(Type = 'Scrambled')

scramble.comparison.object <- rbind(obj.real, obj.scramble)

scramble.comparison.object %>% group_by(Type, VMSgroup) %>% summarise(meanscramble = mean(PC, na.rm=T),
                                                                      sdscramble = sd(PC, na.rm=T))

#shapiro.test(scramble.comparison.object$PC) # not normal


wilcox.test(subset(obj.real, VMSgroup == "withVMS")$PC, 
            subset(obj.scramble, VMSgroup == "withVMS")$PC, conf.int = T) # p<.001

cliff.delta(d = subset(obj.real, VMSgroup == "withVMS")$PC, 
     f = subset(obj.scramble, VMSgroup == "withVMS")$PC)   # .61 large

wilcox.test(subset(obj.real, VMSgroup == "noVMS")$PC, 
            subset(obj.scramble, VMSgroup == "noVMS")$PC, conf.int = T) # ns

cliff.delta(d = subset(obj.real, VMSgroup == "noVMS")$PC, 
     f = subset(obj.scramble, VMSgroup == "noVMS")$PC)   # <.01 negligible

```

```{r Figure 3b}

Figure3b <- 
  ggplot(scramble.comparison.object, aes(x=Type, y=PC, shape = Type, color = VMSgroup))+
  geom_point(aes(group=VMSgroup), size = 5, position=position_dodge(-.2)) +
  stat_summary(fun.data=mean_cl_boot, geom = "pointrange", aes(shape=Type, fill = VMSgroup), 
               size=1.5, position = position_dodge(-.2), colour = "grey32", show.legend = FALSE) +
  stat_summary(fun.y = mean, aes(group=VMSgroup, colour=VMSgroup), geom='line', size=.8, position = position_dodge(-.2), , show.legend = FALSE) +
scale_shape_manual(name = "VMSgroup",
                      labels = c("no-VMS", "with-VMS"),
                       values=c(21,24)) +
    scale_colour_manual(name = "VMSgroup",
                     labels = c("no-VMS", "with-VMS"),
                      values = c("#F8766D","#00BFC4")) +
    xlab("Data type") +
  ylab('% Congruent CPs') +
  ggtitle("Attended Object") +
  theme_bw(base_size=15) +
   theme(axis.text = element_text(size = 20),
         axis.title = element_text(size=20),
         legend.title = element_blank(),
         legend.justification=c(1,1), legend.position=c(1,.99)) +
  guides(shape = FALSE,
         color = guide_legend(override.aes = list(shape=15)))
  
#plot(Figure3b)
multiplot(Figure3a, Figure3b, cols=2)
```


Similar to the caregiver prompts, infants produced more object-matching CPs when the object’s label provided a phonological match with their VMS inventory; this difference was marginally significant (Est.Diff.=.32, p=.056). 
```{r Object inREP vs outREP comparison}

#shapiro.test(subset(match.data.Prompt, VMSgroup=="withVMS")$PC) # normal

match.data.Object %>% 
  filter(VMSgroup == "withVMS") %>% 
  group_by(Type) %>% 
  summarise(meanmatch = mean(PC, na.rm = T),
            sdmatch = sd(PC, na.rm = T))
# 
# t.test(subset(match.data.Object.spread, VMSgroup=="withVMS")$inREP,
# subset(match.data.Object.spread, VMSgroup=="withVMS")$outREP, conf.int=T, paired = T) # p=.052

wilcox.test(subset(match.data.Object.spread, VMSgroup=="withVMS")$inREP,
subset(match.data.Object.spread, VMSgroup=="withVMS")$outREP, conf.int=T, paired = T) # p=.056, Est Diff = .32

wilcoxonPairedR(x = subset(match.data.Object, VMSgroup == "withVMS")$PC,
                g = subset(match.data.Object, VMSgroup == "withVMS")$Type) # r= .42

```

```{r Figure 4b}

Figure4b <- ggplot(match.data.Object, aes(x=Type, y=PC)) +
  geom_point(aes(colour=VMSgroup, shape=Type), size=5, position = position_dodge(-.2)) +
  geom_line(aes(group=subj), colour='grey63',position = position_nudge(x = -.05), linetype=6) +
  stat_summary(fun.data=mean_cl_boot, geom = "pointrange", aes(shape=Type, fill= VMSgroup), colour = "gray32", size=1.5, position = position_dodge(-.2), show.legend = FALSE) +
  stat_summary(fun.y = mean, aes(group=VMSgroup, colour=VMSgroup), geom='line', size=.8, position = position_dodge(-.2), show.legend = FALSE) +
  scale_shape_manual(values=c(23,25)) +
  ylab("% Congruent CPs") +
  xlab("Consonant Type") +
  ggtitle("Attended Object") +
  scale_colour_discrete(breaks=c("noVMS", "withVMS"),
                      labels = c("no-VMS", "with-VMS")) +
  theme_bw(base_size=15) +
   theme(axis.text.x = element_text(size = 20),
         axis.title.x = element_text(size=20),
         axis.title.y = element_text(size=20),
         axis.text.y = element_text(size=20),
                  legend.title = element_blank(),
         legend.justification=c(1,1), legend.position=c(1,.99)) +
  guides(shape = FALSE,
         color = guide_legend(override.aes = list(shape=15)))

multiplot(Figure4a, Figure4b, cols=2)

#plot(Figure4b)
```

Overall, 31% of OUTREP CPs matched an attended object (SD=.29), and this did not differ between VMS groups (Est.Diff.=-.08, p=.32; withVMS: M=.36, SD=.32; noVMS: M=.25, SD=.22). 
```{r Object outREP comparison }

match.data.Object %>% filter(Type == "outREP") %>% summarise(meanMatch = mean(PC, na.rm=T),
                                sdMatch = sd(PC, na.rm=T))

#t.test(PC ~ VMSgroup, subset(match.data.Object, Type=='outREP'), conf.int=T) # ns

wilcox.test(subset(match.data.Object, VMSgroup == "withVMS" & Type == "outREP")$PC, 
            subset(match.data.Object, VMSgroup == "noVMS" & Type == "outREP")$PC, conf.int = T) # ns

cliff.delta(d = subset(match.data.Object, VMSgroup == "withVMS" & Type == "outREP")$PC, 
            f = subset(match.data.Object, VMSgroup == "noVMS" & Type == "outREP")$PC) # .13 negligible

match.data.Object %>% filter(Type == "outREP") %>% group_by(VMSgroup) %>% 
  summarise(meanMatch = mean(PC, na.rm=T),
            sdMatch = sd(PC, na.rm=T))
```




# Discussion

```{r CG responsiveness}

CPdata %>% filter(cgresponse == "None" | is.na(cgresponse)) %>% tally() #1404
1404/1916 #73%

```