---
title: "Supplementary Analyses"
author: "Catherine Laing"
date: "8 March 2019"
output: html_document
---

## Supplementary data

### Part 1: reanalyse variables with matching stimuli removed

# Input prompt

```{r Input prompt stats - reanalysis}

source("ProductionData_nostimmatch_randsubj.R")
source("Data_scrambling_nostimmatch_randsubj.R")
library(effsize)
library(tidyverse)
library(stringi)
library(rcompanion)


1916-1772 # Total CPs in original data - total CPs in data with matches removed = 144

144/1916 #.08 of all CPs from original analysis

```

```{r Re-run analysis with these data removed}

ProdData_nostimmatch %>% ungroup() %>% summarise(totalprompt = sum(nPrompt)) # 839

#shapiro.test(ProdData_nostimmatch$Prod.Prompt.match.PC) # normal

#t.test(Prod.Prompt.match.PC ~ VMSgroup, ProdData_nostimmatch, conf.int=T) # p=.07

wilcox.test(subset(ProdData_nostimmatch, VMSgroup == "withVMS")$Prod.Prompt.match.PC, 
            subset(ProdData_nostimmatch, VMSgroup == "noVMS")$Prod.Prompt.match.PC, conf.int = T) # Significant, p<.001

cliff.delta(d = subset(ProdData_nostimmatch, VMSgroup == "withVMS")$Prod.Prompt.match.PC, 
            f = subset(ProdData_nostimmatch, VMSgroup == "noVMS")$Prod.Prompt.match.PC) # .34, medium

prompt.real <- ProdData_nostimmatch %>%
  select(subj, month, VMSgroup, Prod.Prompt.match.PC) %>%
  ungroup() %>%
  rename(PC = Prod.Prompt.match.PC) %>%
  mutate(Type = 'real')

prompt.scramble <- data.scrambled_nostimmatch %>%
  ungroup() %>%
  select(subj, month, VMSgroup, Prod.Prompt.match.PC) %>%
  rename(PC = Prod.Prompt.match.PC) %>%
  mutate(Type = 'scrambled')

scramble.comparison.prompt <- rbind(prompt.real, prompt.scramble)

wilcox.test(subset(prompt.real, VMSgroup == "withVMS")$PC, 
            subset(prompt.scramble, VMSgroup == "withVMS")$PC, conf.int = T) # Significant, p<.001

cliff.delta(d = subset(prompt.real, VMSgroup == "withVMS")$PC, 
            f = subset(prompt.scramble, VMSgroup == "withVMS")$PC) # .91 large

wilcox.test(subset(prompt.real, VMSgroup == "noVMS")$PC, 
            subset(prompt.scramble, VMSgroup == "noVMS")$PC, conf.int = T) # Significant, p<.05

cliff.delta(d = subset(prompt.real, VMSgroup == "noVMS")$PC, 
            f = subset(prompt.scramble, VMSgroup == "noVMS")$PC) # .40 medium
```

```{r comparing inREP vs outREP consonants - prompt nostimmatch}

#shapiro.test(subset(match.data.Prompt, VMSgroup=="withVMS")$PC) # normal 

# t.test(subset(match.data.Prompt.spread, VMSgroup=="withVMS")$inREP,
# subset(match.data.Prompt.spread, VMSgroup=="withVMS")$outREP, conf.int=T, paired = T) # p<.001

wilcox.test(subset(match.data.Prompt.spread, VMSgroup=="withVMS")$inREP,
subset(match.data.Prompt.spread, VMSgroup=="withVMS")$outREP, conf.int=T, paired = T) # p<.01 EstDiff=.31

wilcoxonPairedR(x = subset(match.data.Prompt, VMSgroup=="withVMS")$PC, 
                g = subset(match.data.Prompt, VMSgroup=="withVMS")$Type) # r = .61

```



# Attended object

```{r Att Obj stats - reanalysis}

ProdData_nostimmatch %>% ungroup() %>% summarise(totalobj = sum(nObj)) # 963

#shapiro.test(ProdData_nostimmatch$Prod.Obj.match.PC) # not normal

wilcox.test(subset(ProdData_nostimmatch, VMSgroup == "withVMS")$Prod.Obj.match.PC, 
            subset(ProdData_nostimmatch, VMSgroup == "noVMS")$Prod.Obj.match.PC, conf.int = T) # p<.01

cliff.delta(d = subset(ProdData_nostimmatch, VMSgroup == "withVMS")$Prod.Obj.match.PC, 
            f = subset(ProdData_nostimmatch, VMSgroup == "noVMS")$Prod.Obj.match.PC) # .47 medium

object.real <- ProdData_nostimmatch %>%
  select(subj, month, VMSgroup, Prod.Obj.match.PC) %>%
  ungroup() %>%
  rename(PC = Prod.Obj.match.PC) %>%
  mutate(Type = 'real')

object.scramble <- data.scrambled_nostimmatch %>%
  ungroup() %>%
  select(subj, month, VMSgroup, Prod.Obj.match.PC) %>%
  rename(PC = Prod.Obj.match.PC) %>%
  mutate(Type = 'scrambled')

scramble.comparison.object <- rbind(object.real, object.scramble)

wilcox.test(subset(object.real, VMSgroup == "withVMS")$PC, 
            subset(object.scramble, VMSgroup == "withVMS")$PC, conf.int = T) # Significant, p<.001

cliff.delta(d = subset(object.real, VMSgroup == "withVMS")$PC, 
            f = subset(object.scramble, VMSgroup == "withVMS")$PC) # .60 large

wilcox.test(subset(object.real, VMSgroup == "noVMS")$PC, 
            subset(object.scramble, VMSgroup == "noVMS")$PC, conf.int = T) # Significant, p<.05

cliff.delta(d = subset(object.real, VMSgroup == "noVMS")$PC, 
            f = subset(object.scramble, VMSgroup == "noVMS")$PC) # -.13 negligible

```

```{r  comparing inREP vs outREP consonants - object nostimmatch}

#shapiro.test(subset(match.data.Object, VMSgroup=="withVMS")$PC) # not normal

wilcox.test(subset(match.data.Object.spread, VMSgroup=="withVMS")$inREP, 
            subset(match.data.Object.spread, VMSgroup=="withVMS")$outREP, conf.int=T, paired = T) # p<.01

wilcoxonPairedR(x = subset(match.data.Object, VMSgroup == "withVMS")$PC,
                g = subset(match.data.Object, VMSgroup == "withVMS")$Type) # r = .38

```


## Part 2: Reanalyse variables after removing objects and prompts with matching onset consonants

# Input Prompts

```{r Testing for C1 match}

# How many CPs match C1 of both prompt and object?

CPdata %>% 
  mutate(match = str_detect(C1Obj, fixed(C1Prompt)),  # Is there a consonantal match: T/F
         match = ifelse(C1Obj == "none" & C1Prompt == "n", F, match),      
         # ^ none matches n in above line, so make sure these are included here
         match = ifelse(C1Prompt == "none" & C1Obj == "n", F, match)) %>% # and vice versa
  filter(match == TRUE & C1Obj != "none") %>% tally() # 223    # exclude all matches of 'none'

223/1916 # 12% of original dataset

source("ProductionData_noC1match_randsubj.R")
source("Data_scrambling_noC1match_randsubj.R")

1916-1693 # 223   # check that numbers are correct in new CP data 

```

```{r Testing with noC1 match - Prompt}

ProdData_noC1match %>% ungroup() %>% summarise(totalprompt = sum(nPrompt)) # 753

#shapiro.test(ProdData_noC1match$Prod.Prompt.match.PC) # normal

# t.test(Prod.Prompt.match.PC ~ VMSgroup, ProdData_noC1match, conf.int=T) # ns

wilcox.test(subset(ProdData_noC1match, VMSgroup == "withVMS")$Prod.Prompt.match.PC, 
            subset(ProdData_noC1match, VMSgroup == "noVMS")$Prod.Prompt.match.PC, conf.int = T) # ns

cliff.delta(d = subset(ProdData_noC1match, VMSgroup == "withVMS")$Prod.Prompt.match.PC, 
            f = subset(ProdData_noC1match, VMSgroup == "noVMS")$Prod.Prompt.match.PC) # .26 small

prompt.real <- ProdData_noC1match %>%
  select(subj, month, VMSgroup, Prod.Prompt.match.PC) %>%
  ungroup() %>%
  rename(PC = Prod.Prompt.match.PC) %>%
  mutate(Type = 'real')

prompt.scramble <- data.scrambled_noC1match %>%
  ungroup() %>%
  select(subj, month, VMSgroup, Prod.Prompt.match.PC) %>%
  rename(PC = Prod.Prompt.match.PC) %>%
  mutate(Type = 'scrambled')

scramble.comparison.prompt <- rbind(prompt.real, prompt.scramble)

wilcox.test(subset(prompt.real, VMSgroup == "withVMS")$PC, 
            subset(prompt.scramble, VMSgroup == "withVMS")$PC, conf.int = T) # Significant, p<.001

cliff.delta(d = subset(prompt.real, VMSgroup == "withVMS")$PC, 
            f = subset(prompt.scramble, VMSgroup == "withVMS")$PC) # Significant, p<.001

wilcox.test(subset(prompt.real, VMSgroup == "noVMS")$PC, 
            subset(prompt.scramble, VMSgroup == "noVMS")$PC, conf.int = T) # Significant, p<.05

cliff.delta(d = subset(prompt.real, VMSgroup == "noVMS")$PC, 
            f = subset(prompt.scramble, VMSgroup == "noVMS")$PC) # Significant, p<.05

```

```{r comparing inREP vs outREP consonants - Prompt noC1match}

#shapiro.test(match.data.Prompt$PC) # not normal

wilcox.test(subset(match.data.Prompt.spread, VMSgroup=="withVMS")$inREP, 
            subset(match.data.Prompt.spread, VMSgroup=="withVMS")$outREP, conf.int=T, paired = T) # p<.01

wilcoxonPairedR(x = subset(match.data.Prompt, VMSgroup == "withVMS")$PC,
                g = subset(match.data.Prompt, VMSgroup == "withVMS")$Type) # r = .61
```

# Attended Object

```{r r Testing with noC1 match - Object}

ProdData_noC1match %>% ungroup() %>% summarise(totalobj = sum(nObj)) # 877

#shapiro.test(ProdData_noC1match$Prod.Obj.match.PC) # not normal

wilcox.test(subset(ProdData_noC1match, VMSgroup == "withVMS")$Prod.Obj.match.PC, 
            subset(ProdData_noC1match, VMSgroup == "noVMS")$Prod.Obj.match.PC, conf.int = T) # p<.01

cliff.delta(d = subset(ProdData_noC1match, VMSgroup == "withVMS")$Prod.Obj.match.PC, 
            f = subset(ProdData_noC1match, VMSgroup == "noVMS")$Prod.Obj.match.PC) # .48 large

object.real <- ProdData_noC1match %>%
  select(subj, month, VMSgroup, Prod.Obj.match.PC) %>%
  ungroup() %>%
  rename(PC = Prod.Obj.match.PC) %>%
  mutate(Type = 'real')

object.scramble <- data.scrambled_noC1match %>%
  ungroup() %>%
  select(subj, month, VMSgroup, Prod.Obj.match.PC) %>%
  rename(PC = Prod.Obj.match.PC) %>%
  mutate(Type = 'scrambled')

scramble.comparison.object <- rbind(object.real, object.scramble)

wilcox.test(subset(object.real, VMSgroup == "withVMS")$PC, 
            subset(object.scramble, VMSgroup == "withVMS")$PC, conf.int = T) # Significant, p<.001

cliff.delta(d = subset(object.real, VMSgroup == "withVMS")$PC, 
            f = subset(object.scramble, VMSgroup == "withVMS")$PC) # .60 large

wilcox.test(subset(object.real, VMSgroup == "noVMS")$PC, 
            subset(object.scramble, VMSgroup == "noVMS")$PC, conf.int = T) # ns

cliff.delta(d = subset(object.real, VMSgroup == "noVMS")$PC, 
            f = subset(object.scramble, VMSgroup == "noVMS")$PC) # -.11 negligible
```

```{r comparing inREP vs outREP consonants - Object noC1match}

#shapiro.test(subset(match.data.Object, VMSgroup=="withVMS")$PC) # not normal

wilcox.test(subset(match.data.Object.spread, VMSgroup=="withVMS")$inREP, 
            subset(match.data.Object.spread, VMSgroup=="withVMS")$outREP, conf.int=T, paired = T) # p=.07

wilcoxonPairedR(x = subset(match.data.Object, VMSgroup == "withVMS")$PC,
                g = subset(match.data.Object, VMSgroup == "withVMS")$Type) # r = .61

```


## Part 3: Reanlalyse caregiver input with a subset of data that only includes nouns, onomtopoeia and routine words in CPdata$cgprompt

```{r Source data}

source("Data_scrambling_Nounsonly_randsubj.R")

```

```{r Overview}

# 752 CPs with noun/onomatopoeia as prompts

CPdata_nouns %>% group_by(wordclass) %>% tally()  #32 onomatopoeia, 39 greetings/routine words
                              # (cf Laing et al 2015 who show that onomatopoiea are particularly salient in caregier input)

```

```{r Re-run analysis}

#shapiro.test(ProdData_nouns$Prod.Prompt.match.PC) # normal

#t.test(Prod.Prompt.match.PC ~ VMSgroup, ProdData_nouns) # p=.11

wilcox.test(subset(ProdData_nouns, VMSgroup == "withVMS")$Prod.Prompt.match.PC, 
            subset(ProdData_nouns, VMSgroup == "noVMS")$Prod.Prompt.match.PC, conf.int = T) # p=.17

cliff.delta(d = subset(ProdData_nouns, VMSgroup == "withVMS")$Prod.Prompt.match.PC, 
            f = subset(ProdData_nouns, VMSgroup == "noVMS")$Prod.Prompt.match.PC) # p=.24 small


prompt.real <- ProdData_nouns %>%
  select(subj, month, VMSgroup, Prod.Prompt.match.PC) %>%
  ungroup() %>%
  rename(PC = Prod.Prompt.match.PC) %>%
  mutate(Type = 'real')

prompt.scramble <- data.scrambled %>%
  ungroup() %>%
  select(subj, month, VMSgroup, Prod.Prompt.match.PC) %>%
  rename(PC = Prod.Prompt.match.PC) %>%
  mutate(Type = 'scrambled')

scramble.comparison.prompt <- rbind(prompt.real, prompt.scramble)

#shapiro.test(scramble.comparison.prompt$PC) # not normal

wilcox.test(subset(prompt.real, VMSgroup == "withVMS")$PC, 
            subset(prompt.scramble, VMSgroup == "withVMS")$PC, conf.int = T) # Significant, p<.001

cliff.delta(d = subset(prompt.real, VMSgroup == "withVMS")$PC, 
            f = subset(prompt.scramble, VMSgroup == "withVMS")$PC) # .90 large

wilcox.test(subset(prompt.real, VMSgroup == "noVMS")$PC, 
            subset(prompt.scramble, VMSgroup == "noVMS")$PC, conf.int = T) # Significant, p<.05

cliff.delta(d = subset(prompt.real, VMSgroup == "noVMS")$PC, 
            f = subset(prompt.scramble, VMSgroup == "noVMS")$PC) # .41 medium
```

```{r comparing inREP vs outREP consonants - nouns only}

#shapiro.test(subset(match.data.Prompt, VMSgroup=="withVMS")$PC) # normal 

# t.test(subset(match.data.Prompt_nouns.spread, VMSgroup=="withVMS")$inREP,
# subset(match.data.Prompt_nouns.spread, VMSgroup=="withVMS")$outREP, conf.int=T, paired = T) # p<.01

wilcox.test(subset(match.data.Prompt_nouns.spread, VMSgroup=="withVMS")$inREP, 
            subset(match.data.Prompt_nouns.spread, VMSgroup=="withVMS")$outREP, conf.int=T, paired = T) # p<.01

wilcoxonPairedR(x = subset(match.data.Prompt_nouns, VMSgroup == "withVMS")$PC,
                g = subset(match.data.Prompt_nouns, VMSgroup == "withVMS")$Type) # r = .58

```








