---
title: "Supplementary Analyses: From babble to words"
author: "Catherine Laing & Elika Bergelson"
date: "7 August 2019"
output: html_document
---

## Supplementary data

```{r Set up, include=FALSE}
library(tidyverse)
library(effsize)
library(rcompanion)

```

### Part 1: Running analysis with three infants removed (all produced <50 but >20 tokens of a given consonant at 0;10 and 0;11)

```{r source data}

source("ProductionData_randsubj.R")
source("Data_scrambling_randsubj.R")

```

```{r check audio production}

toexclude <- vmscount %>% 
  group_by(subj, VMSgroup, month_chron, n) %>% 
  filter(subj %in% vms_subset2$subj &
           month_chron == 11 & 
           n < 50 & n >19 &
           inREP == T) %>%
  #view()
  ungroup() %>%
  group_by(subj) %>%
  tally()   # 3 infants, 279, 129, 268

vmscount %>%
  group_by(subj, Consonant, inREP) %>%
  summarise(totalprod = sum(n)) %>%
  filter(subj %in% toexclude$subj & inREP == T & totalprod < 50)  # all consonant production >50 tokens with months combined

```

```{r exclude these infants from datasets ro re-run analysis}

ProdData_subset <- ProdData %>%
  filter(!subj %in% toexclude$subj)

data.scrambled_subset <- data.scrambled %>%
  filter(!subj %in% toexclude$subj)

vmstotal_subset <- vmstotal %>%
  filter(!subj %in% toexclude$subj)

compareRecordings_subset <- compareRecordings %>%
  filter(!subj %in% toexclude$subj)

compareRecordings.spread_subset <- compareRecordings.spread %>%
  filter(!subj %in% toexclude$subj)

match.data.Prompt_subset <- match.data.Prompt %>%
  filter(!subj %in% toexclude$subj)

match.data.Prompt.spread_subset <- match.data.Prompt.spread %>%
  filter(!subj %in% toexclude$subj)

match.data.Object_subset <- match.data.Object %>%
  filter(!subj %in% toexclude$subj)

match.data.Object.spread_subset <- match.data.Object.spread %>%
  filter(!subj %in% toexclude$subj)


```


```{r comparing production across groups in audio recordings - subset}

shapiro.test(vmstotal_subset$audiotokens) # not normal

wilcox.test(subset(vmstotal_subset, VMSgroup == "withVMS")$audiotokens, subset(vmstotal_subset, VMSgroup == "noVMS")$audiotokens, conf.int=T) # p<.001

cliff.delta(d = subset(vmstotal_subset, VMSgroup == "withVMS")$audiotokens, 
     f = subset(vmstotal_subset, VMSgroup == "noVMS")$audiotokens)   # .99 large


shapiro.test(vmstotal_subset$audiotypes) # not normal

wilcox.test(subset(vmstotal_subset, VMSgroup == "withVMS")$audiotypes, subset(vmstotal_subset, VMSgroup == "noVMS")$audiotypes, conf.int=T) # p<.001

cliff.delta(d = subset(vmstotal_subset, VMSgroup == "withVMS")$audiotypes, 
     f = subset(vmstotal_subset, VMSgroup == "noVMS")$audiotypes)   # .74 large

```

# Infant Production: Video recordings

```{r Effect of VMS group on CP types/tokens}

wilcox.test(subset(ProdData_subset, VMSgroup == "withVMS")$CPtokens, subset(ProdData_subset, VMSgroup == "noVMS")$CPtokens, conf.int=T) # p<.001

cliff.delta(d = subset(ProdData_subset, VMSgroup == "withVMS")$CPtokens, 
     f = subset(ProdData_subset, VMSgroup == "noVMS")$CPtokens)   # .68 large

wilcox.test(subset(ProdData_subset, VMSgroup == "withVMS")$CPtypes, subset(ProdData_subset, VMSgroup == "noVMS")$CPtypes, conf.int=T) # p=.15 

cliff.delta(d = subset(ProdData_subset, VMSgroup == "withVMS")$CPtypes, 
     f = subset(ProdData_subset, VMSgroup == "noVMS")$CPtypes)   # .32 small

```

```{r calcuating number of in vs out VMS consonants in video recordings}

shapiro.test(compareRecordings_subset$meanProd) # not normal

wilcox.test(compareRecordings.spread_subset$VMS.Prod.Cong, compareRecordings.spread_subset$VMS.Prod.Incong, paired=T, conf.int=T) # p<.01

wilcoxonPairedR(x = compareRecordings_subset$meanProd,
                g = compareRecordings_subset$VMS.Prod.Cong) # r=-.58

```
## Input prompts

```{r Input prompt overview}

#shapiro.test(ProdData_subset$Prod.Prompt.match.PC) # p=.03

#t.test(Prod.Prompt.match.PC ~ VMSgroup, ProdData_subset) # p=.04

wilcox.test(subset(ProdData_subset, VMSgroup=="withVMS")$Prod.Prompt.match.PC, 
            subset(ProdData_subset, VMSgroup=="noVMS")$Prod.Prompt.match.PC, conf.int=T) # p=.09, EstDiff=-.13

cliff.delta(d = subset(ProdData_subset, VMSgroup=="withVMS")$Prod.Prompt.match.PC, 
     f = subset(ProdData_subset, VMSgroup=="noVMS")$Prod.Prompt.match.PC)   # .32 small

wilcox.test(subset(ProdData_subset, VMSgroup=="withVMS" & !subj %in% c(279, 129, 268))$Prod.Prompt.match.PC, 
            subset(ProdData_subset, VMSgroup=="noVMS")$Prod.Prompt.match.PC, conf.int=T) # p=.09, EstDiff=-.13

```

WithVMS infants’ CPs matched the consonantal properties of their caregivers’ speech significantly more in the real data set than in the scrambled one (.55(.17) vs. .22(.14), Est.Diff.=.33, p<.001 by Wilcoxon Test), as did noVMS infants’ CPs (.41(.28) vs. .2(.16), Est.Diff.=.24, p<.01). 
```{r Input prompt data vs scrambled data}

prompt.real <- ProdData_subset %>%
  select(subj, month, VMSgroup, Prod.Prompt.match.PC) %>%
  ungroup() %>%
  rename(PC = Prod.Prompt.match.PC) %>%
  mutate(Type = 'Real')

prompt.scramble <- data.scrambled_subset %>%
  ungroup() %>%
  select(subj, month, VMSgroup, Prod.Prompt.match.PC) %>%
  rename(PC = Prod.Prompt.match.PC) %>%
  mutate(Type = 'Scrambled')

scramble.comparison.prompt <- rbind(prompt.real, prompt.scramble)

#shapiro.test(scramble.comparison.prompt$PC) # not normal

wilcox.test(subset(prompt.real, VMSgroup == "withVMS")$PC, 
            subset(prompt.scramble, VMSgroup == "withVMS")$PC, conf.int = T) # Significant, p<.001

cliff.delta(d = subset(prompt.real, VMSgroup == "withVMS")$PC, 
     f = subset(prompt.scramble, VMSgroup == "withVMS")$PC)   # .92 large

wilcox.test(subset(prompt.real, VMSgroup == "noVMS")$PC, 
            subset(prompt.scramble, VMSgroup == "noVMS")$PC, conf.int = T) # Significant, p<.05

cliff.delta(d = subset(prompt.real, VMSgroup == "noVMS")$PC, 
     f = subset(prompt.scramble, VMSgroup == "noVMS")$PC)   # .39 medium

```

```{r Input prompt inREP vs outREP comparison}

shapiro.test(subset(match.data.Prompt_subset, VMSgroup=="withVMS")$PC) #normal

# t.test(subset(match.data.Prompt.spread_subset, VMSgroup=="withVMS")$inREP,
# subset(match.data.Prompt.spread_subset, VMSgroup=="withVMS")$outREP, conf.int=T, paired = T) # p=.001

wilcox.test(subset(match.data.Prompt.spread_subset, VMSgroup=="withVMS")$inREP,
            subset(match.data.Prompt.spread_subset, VMSgroup=="withVMS")$outREP, conf.int=T, paired = T) # p<.01, Est.Diff=.34

wilcoxonPairedR(x = subset(match.data.Prompt_subset, VMSgroup == "withVMS")$PC,
                g = subset(match.data.Prompt_subset, VMSgroup == "withVMS")$Type) # r= .64

```
```{r Input prompt outREP comparison}

shapiro.test(subset(match.data.Prompt_subset, Type=='outREP')$PC)#normal

#t.test(PC ~ VMSgroup, subset(match.data.Prompt_subset, Type=='outREP'), conf.int=T) # ns

wilcox.test(subset(match.data.Prompt_subset, VMSgroup == "withVMS" & Type == "outREP")$PC, 
            subset(match.data.Prompt_subset, VMSgroup == "noVMS" & Type == "outREP")$PC, conf.int = T) # ns

cliff.delta(d = subset(match.data.Prompt_subset, VMSgroup == "withVMS" & Type == "outREP")$PC, 
     f = subset(match.data.Prompt_subset, VMSgroup == "noVMS" & Type == "outREP")$PC)   # -.05 negligible

```


## Attended objects

```{r Object overview}

wilcox.test(subset(ProdData_subset, VMSgroup == "withVMS")$Prod.Obj.match.PC, 
            subset(ProdData_subset, VMSgroup == "noVMS")$Prod.Obj.match.PC, conf.int = T) # p<.01

cliff.delta(d = subset(ProdData_subset, VMSgroup == "withVMS")$Prod.Obj.match.PC, 
     f = subset(ProdData, VMSgroup == "noVMS")$Prod.Obj.match.PC)   # .5 large

```


```{r Object data vs scrambled data}

obj.real <- ProdData_subset %>%
  select(subj, month, VMSgroup, Prod.Obj.match.PC) %>%
  ungroup() %>%
  rename(PC = Prod.Obj.match.PC) %>%
  mutate(Type = 'Real')

obj.scramble <- data.scrambled_subset %>%
  ungroup() %>%
  select(subj, month, VMSgroup, Prod.Obj.match.PC) %>%
  rename(PC = Prod.Obj.match.PC) %>%
  mutate(Type = 'Scrambled')

scramble.comparison.object <- rbind(obj.real, obj.scramble)

#shapiro.test(scramble.comparison.object$PC) # not normal

wilcox.test(subset(obj.real, VMSgroup == "withVMS")$PC, 
            subset(obj.scramble, VMSgroup == "withVMS")$PC, conf.int = T) # p<.001

cliff.delta(d = subset(obj.real, VMSgroup == "withVMS")$PC, 
     f = subset(obj.scramble, VMSgroup == "withVMS")$PC)   # .67 large

wilcox.test(subset(obj.real, VMSgroup == "noVMS")$PC, 
            subset(obj.scramble, VMSgroup == "noVMS")$PC, conf.int = T) # ns

cliff.delta(d = subset(obj.real, VMSgroup == "noVMS")$PC, 
     f = subset(obj.scramble, VMSgroup == "noVMS")$PC)   # <-.1 negligible

```


```{r Object inREP vs outREP comparison}

# shapiro.test(subset(match.data.Object_subset, VMSgroup=="withVMS")$PC) # not normal

wilcox.test(subset(match.data.Object.spread_subset, VMSgroup=="withVMS")$inREP, 
            subset(match.data.Object.spread_subset, VMSgroup=="withVMS")$outREP, conf.int=T, paired = T) # p=.11, Est Diff = .31

wilcoxonPairedR(x = subset(match.data.Object_subset, VMSgroup == "withVMS")$PC,
                g = subset(match.data.Object_subset, VMSgroup == "withVMS")$Type) # r= .38

```


```{r Object outREP comparison }

wilcox.test(subset(match.data.Object_subset, VMSgroup == "withVMS" & Type == "outREP")$PC, 
            subset(match.data.Object_subset, VMSgroup == "noVMS" & Type == "outREP")$PC, conf.int = T) # ns


cliff.delta(d = subset(match.data.Object_subset, VMSgroup == "withVMS" & Type == "outREP")$PC, 
            f = subset(match.data.Object_subset, VMSgroup == "noVMS" & Type == "outREP")$PC) # .17 small

```


### Part 2: Figure S5

```{r data and theme, include=FALSE}

mytheme <- theme(strip.text.x = element_blank(),
      axis.text.x  = element_blank(),
      axis.title.y = element_text(),
      axis.text.y = element_text(size=7),
      axis.ticks.x = element_blank(),
      legend.text = element_text(),
      legend.title = element_text(),
      panel.grid = element_blank(),
      panel.border = element_blank(),
      panel.spacing.x = unit(0.,"lines")
)

```

```{r just10mo, echo = F}


figS5a <- ggplot(vmscount 
                %>% filter(month==10 & month_chron==10), 
                aes(subj,
                    y=n, fill = Consonant,
                    linetype=VMSgroup))  +  
  geom_col(lwd = .3, color = "black") +
  facet_grid(month_chron~subj, scales = "free")+
  scale_fill_discrete(breaks = c("p", "t", "k", "m", "n")) +
  xlab("Individual Subjects") + 
  ylab("n tokens") +
  scale_y_continuous(breaks = c(0, 20, 50, 200, 400)) +
  scale_linetype_manual(values = c(0,1))+
  theme_bw()+
  mytheme+
  labs(linetype = "VMS") +
  geom_hline(yintercept = 20, linetype = 2, colour = "grey27") +
  geom_hline(yintercept = 50, linetype = 4, colour = "grey27")

plot(figS5a)


```

```{r month10and11, echo = F}
figS5b <- ggplot(vmscount 
       %>% filter(month==11), 
       aes(subj,
           y=n, fill = Consonant,
           linetype=VMSgroup))  +  
  geom_col(lwd = .3, color = "black") +
  facet_grid(month_chron~subj, scales = "free")+
  scale_fill_discrete(breaks = c("p", "t", "k", "m", "n"), guide = F) +
  xlab("Individual Subjects") + 
  ylab("n tokens") +
  scale_y_continuous(breaks = c(0, 20, 50, 200, 400)) +
  scale_linetype_manual(values = c(0,1))+
  theme_bw()+
  mytheme+ 
  labs(linetype = "VMS") +
  geom_hline(yintercept = 20, linetype = 2, colour = "grey27") +
  geom_hline(yintercept = 50, linetype = 4, colour = "grey27")

plot(figS5b)
```

### Part 3: reanalyse variables with matching stimuli removed

# Input prompt

```{r Input prompt stats - reanalysis}

source("ProductionData_nostimmatch_randsubj.R")
source("Data_scrambling_nostimmatch_randsubj.R")

1916-1772 # Total CPs in original data - total CPs in data with matches removed = 144

144/1916 #.08 of all CPs from original analysis

```

```{r Re-run analysis with these data removed}

ProdData_nostimmatch %>% ungroup() %>% summarise(totalprompt = sum(nPrompt)) # 839

#shapiro.test(ProdData_nostimmatch$Prod.Prompt.match.PC) # normal

#t.test(Prod.Prompt.match.PC ~ VMSgroup, ProdData_nostimmatch, conf.int=T) # p=.07

wilcox.test(subset(ProdData_nostimmatch, VMSgroup == "withVMS")$Prod.Prompt.match.PC, 
            subset(ProdData_nostimmatch, VMSgroup == "noVMS")$Prod.Prompt.match.PC, conf.int = T) # Significant, p<.001

cliff.delta(d = subset(ProdData_nostimmatch, VMSgroup == "withVMS")$Prod.Prompt.match.PC, 
            f = subset(ProdData_nostimmatch, VMSgroup == "noVMS")$Prod.Prompt.match.PC) # .34, medium

prompt.real <- ProdData_nostimmatch %>%
  select(subj, month, VMSgroup, Prod.Prompt.match.PC) %>%
  ungroup() %>%
  rename(PC = Prod.Prompt.match.PC) %>%
  mutate(Type = 'real')

prompt.scramble <- data.scrambled_nostimmatch %>%
  ungroup() %>%
  select(subj, month, VMSgroup, Prod.Prompt.match.PC) %>%
  rename(PC = Prod.Prompt.match.PC) %>%
  mutate(Type = 'scrambled')

scramble.comparison.prompt <- rbind(prompt.real, prompt.scramble)

wilcox.test(subset(prompt.real, VMSgroup == "withVMS")$PC, 
            subset(prompt.scramble, VMSgroup == "withVMS")$PC, conf.int = T) # Significant, p<.001

cliff.delta(d = subset(prompt.real, VMSgroup == "withVMS")$PC, 
            f = subset(prompt.scramble, VMSgroup == "withVMS")$PC) # .91 large

wilcox.test(subset(prompt.real, VMSgroup == "noVMS")$PC, 
            subset(prompt.scramble, VMSgroup == "noVMS")$PC, conf.int = T) # Significant, p<.05

cliff.delta(d = subset(prompt.real, VMSgroup == "noVMS")$PC, 
            f = subset(prompt.scramble, VMSgroup == "noVMS")$PC) # .40 medium
```

```{r comparing inREP vs outREP consonants - prompt nostimmatch}

#shapiro.test(subset(match.data.Prompt, VMSgroup=="withVMS")$PC) # normal 

# t.test(subset(match.data.Prompt.spread, VMSgroup=="withVMS")$inREP,
# subset(match.data.Prompt.spread, VMSgroup=="withVMS")$outREP, conf.int=T, paired = T) # p<.001

wilcox.test(subset(match.data.Prompt.spread, VMSgroup=="withVMS")$inREP,
subset(match.data.Prompt.spread, VMSgroup=="withVMS")$outREP, conf.int=T, paired = T) # p<.01 EstDiff=.31

wilcoxonPairedR(x = subset(match.data.Prompt, VMSgroup=="withVMS")$PC, 
                g = subset(match.data.Prompt, VMSgroup=="withVMS")$Type) # r = .61

```



# Attended object

```{r Att Obj stats - reanalysis}

ProdData_nostimmatch %>% ungroup() %>% summarise(totalobj = sum(nObj)) # 963

#shapiro.test(ProdData_nostimmatch$Prod.Obj.match.PC) # not normal

wilcox.test(subset(ProdData_nostimmatch, VMSgroup == "withVMS")$Prod.Obj.match.PC, 
            subset(ProdData_nostimmatch, VMSgroup == "noVMS")$Prod.Obj.match.PC, conf.int = T) # p<.01

cliff.delta(d = subset(ProdData_nostimmatch, VMSgroup == "withVMS")$Prod.Obj.match.PC, 
            f = subset(ProdData_nostimmatch, VMSgroup == "noVMS")$Prod.Obj.match.PC) # .47 medium

object.real <- ProdData_nostimmatch %>%
  select(subj, month, VMSgroup, Prod.Obj.match.PC) %>%
  ungroup() %>%
  rename(PC = Prod.Obj.match.PC) %>%
  mutate(Type = 'real')

object.scramble <- data.scrambled_nostimmatch %>%
  ungroup() %>%
  select(subj, month, VMSgroup, Prod.Obj.match.PC) %>%
  rename(PC = Prod.Obj.match.PC) %>%
  mutate(Type = 'scrambled')

scramble.comparison.object <- rbind(object.real, object.scramble)

wilcox.test(subset(object.real, VMSgroup == "withVMS")$PC, 
            subset(object.scramble, VMSgroup == "withVMS")$PC, conf.int = T) # Significant, p<.001

cliff.delta(d = subset(object.real, VMSgroup == "withVMS")$PC, 
            f = subset(object.scramble, VMSgroup == "withVMS")$PC) # .60 large

wilcox.test(subset(object.real, VMSgroup == "noVMS")$PC, 
            subset(object.scramble, VMSgroup == "noVMS")$PC, conf.int = T) # Significant, p<.05

cliff.delta(d = subset(object.real, VMSgroup == "noVMS")$PC, 
            f = subset(object.scramble, VMSgroup == "noVMS")$PC) # -.13 negligible

```

```{r  comparing inREP vs outREP consonants - object nostimmatch}

#shapiro.test(subset(match.data.Object, VMSgroup=="withVMS")$PC) # not normal

wilcox.test(subset(match.data.Object.spread, VMSgroup=="withVMS")$inREP, 
            subset(match.data.Object.spread, VMSgroup=="withVMS")$outREP, conf.int=T, paired = T) # p<.01

wilcoxonPairedR(x = subset(match.data.Object, VMSgroup == "withVMS")$PC,
                g = subset(match.data.Object, VMSgroup == "withVMS")$Type) # r = .38

```


## Part 4: Reanalyse variables after removing objects and prompts with matching onset consonants

# Input Prompts

```{r Testing for C1 match}

# How many CPs match C1 of both prompt and object?

CPdata %>% 
  mutate(match = str_detect(C1Obj, fixed(C1Prompt)),  # Is there a consonantal match: T/F
         match = ifelse(C1Obj == "none" & C1Prompt == "n", F, match),      
         # ^ none matches n in above line, so make sure these are included here
         match = ifelse(C1Prompt == "none" & C1Obj == "n", F, match)) %>% # and vice versa
  filter(match == TRUE & C1Obj != "none") %>% tally() # 223    # exclude all matches of 'none'

223/1916 # 12% of original dataset

source("ProductionData_noC1match_randsubj.R")
source("Data_scrambling_noC1match_randsubj.R")

1916-1693 # 223   # check that numbers are correct in new CP data 

```

```{r Testing with noC1 match - Prompt}

ProdData_noC1match %>% ungroup() %>% summarise(totalprompt = sum(nPrompt)) # 753

#shapiro.test(ProdData_noC1match$Prod.Prompt.match.PC) # normal

# t.test(Prod.Prompt.match.PC ~ VMSgroup, ProdData_noC1match, conf.int=T) # ns

wilcox.test(subset(ProdData_noC1match, VMSgroup == "withVMS")$Prod.Prompt.match.PC, 
            subset(ProdData_noC1match, VMSgroup == "noVMS")$Prod.Prompt.match.PC, conf.int = T) # ns

cliff.delta(d = subset(ProdData_noC1match, VMSgroup == "withVMS")$Prod.Prompt.match.PC, 
            f = subset(ProdData_noC1match, VMSgroup == "noVMS")$Prod.Prompt.match.PC) # .26 small

prompt.real <- ProdData_noC1match %>%
  select(subj, month, VMSgroup, Prod.Prompt.match.PC) %>%
  ungroup() %>%
  rename(PC = Prod.Prompt.match.PC) %>%
  mutate(Type = 'real')

prompt.scramble <- data.scrambled_noC1match %>%
  ungroup() %>%
  select(subj, month, VMSgroup, Prod.Prompt.match.PC) %>%
  rename(PC = Prod.Prompt.match.PC) %>%
  mutate(Type = 'scrambled')

scramble.comparison.prompt <- rbind(prompt.real, prompt.scramble)

wilcox.test(subset(prompt.real, VMSgroup == "withVMS")$PC, 
            subset(prompt.scramble, VMSgroup == "withVMS")$PC, conf.int = T) # Significant, p<.001

cliff.delta(d = subset(prompt.real, VMSgroup == "withVMS")$PC, 
            f = subset(prompt.scramble, VMSgroup == "withVMS")$PC) # Significant, p<.001

wilcox.test(subset(prompt.real, VMSgroup == "noVMS")$PC, 
            subset(prompt.scramble, VMSgroup == "noVMS")$PC, conf.int = T) # Significant, p<.05

cliff.delta(d = subset(prompt.real, VMSgroup == "noVMS")$PC, 
            f = subset(prompt.scramble, VMSgroup == "noVMS")$PC) # Significant, p<.05

```

```{r comparing inREP vs outREP consonants - Prompt noC1match}

#shapiro.test(match.data.Prompt$PC) # not normal

wilcox.test(subset(match.data.Prompt.spread, VMSgroup=="withVMS")$inREP, 
            subset(match.data.Prompt.spread, VMSgroup=="withVMS")$outREP, conf.int=T, paired = T) # p<.01

wilcoxonPairedR(x = subset(match.data.Prompt, VMSgroup == "withVMS")$PC,
                g = subset(match.data.Prompt, VMSgroup == "withVMS")$Type) # r = .61
```

# Attended Object

```{r r Testing with noC1 match - Object}

ProdData_noC1match %>% ungroup() %>% summarise(totalobj = sum(nObj)) # 877

#shapiro.test(ProdData_noC1match$Prod.Obj.match.PC) # not normal

wilcox.test(subset(ProdData_noC1match, VMSgroup == "withVMS")$Prod.Obj.match.PC, 
            subset(ProdData_noC1match, VMSgroup == "noVMS")$Prod.Obj.match.PC, conf.int = T) # p<.01

cliff.delta(d = subset(ProdData_noC1match, VMSgroup == "withVMS")$Prod.Obj.match.PC, 
            f = subset(ProdData_noC1match, VMSgroup == "noVMS")$Prod.Obj.match.PC) # .48 large

object.real <- ProdData_noC1match %>%
  select(subj, month, VMSgroup, Prod.Obj.match.PC) %>%
  ungroup() %>%
  rename(PC = Prod.Obj.match.PC) %>%
  mutate(Type = 'real')

object.scramble <- data.scrambled_noC1match %>%
  ungroup() %>%
  select(subj, month, VMSgroup, Prod.Obj.match.PC) %>%
  rename(PC = Prod.Obj.match.PC) %>%
  mutate(Type = 'scrambled')

scramble.comparison.object <- rbind(object.real, object.scramble)

wilcox.test(subset(object.real, VMSgroup == "withVMS")$PC, 
            subset(object.scramble, VMSgroup == "withVMS")$PC, conf.int = T) # Significant, p<.001

cliff.delta(d = subset(object.real, VMSgroup == "withVMS")$PC, 
            f = subset(object.scramble, VMSgroup == "withVMS")$PC) # .60 large

wilcox.test(subset(object.real, VMSgroup == "noVMS")$PC, 
            subset(object.scramble, VMSgroup == "noVMS")$PC, conf.int = T) # ns

cliff.delta(d = subset(object.real, VMSgroup == "noVMS")$PC, 
            f = subset(object.scramble, VMSgroup == "noVMS")$PC) # -.11 negligible
```

```{r comparing inREP vs outREP consonants - Object noC1match}

#shapiro.test(subset(match.data.Object, VMSgroup=="withVMS")$PC) # not normal

wilcox.test(subset(match.data.Object.spread, VMSgroup=="withVMS")$inREP, 
            subset(match.data.Object.spread, VMSgroup=="withVMS")$outREP, conf.int=T, paired = T) # p=.07

wilcoxonPairedR(x = subset(match.data.Object, VMSgroup == "withVMS")$PC,
                g = subset(match.data.Object, VMSgroup == "withVMS")$Type) # r = .61

```


## Part 3: Reanlalyse caregiver input with a subset of data that only includes nouns, onomtopoeia and routine words in CPdata$cgprompt

```{r Source data}

source("Data_scrambling_Nounsonly_randsubj.R")

```

```{r Overview}

# 752 CPs with noun/onomatopoeia as prompts

CPdata_nouns %>% group_by(wordclass) %>% tally()  #32 onomatopoeia, 39 greetings/routine words
                              # (cf Laing et al 2015 who show that onomatopoiea are particularly salient in caregier input)

```

```{r Re-run analysis}

#shapiro.test(ProdData_nouns$Prod.Prompt.match.PC) # normal

#t.test(Prod.Prompt.match.PC ~ VMSgroup, ProdData_nouns) # p=.11

wilcox.test(subset(ProdData_nouns, VMSgroup == "withVMS")$Prod.Prompt.match.PC, 
            subset(ProdData_nouns, VMSgroup == "noVMS")$Prod.Prompt.match.PC, conf.int = T) # p=.17

cliff.delta(d = subset(ProdData_nouns, VMSgroup == "withVMS")$Prod.Prompt.match.PC, 
            f = subset(ProdData_nouns, VMSgroup == "noVMS")$Prod.Prompt.match.PC) # p=.24 small


prompt.real <- ProdData_nouns %>%
  select(subj, month, VMSgroup, Prod.Prompt.match.PC) %>%
  ungroup() %>%
  rename(PC = Prod.Prompt.match.PC) %>%
  mutate(Type = 'real')

prompt.scramble <- data.scrambled %>%
  ungroup() %>%
  select(subj, month, VMSgroup, Prod.Prompt.match.PC) %>%
  rename(PC = Prod.Prompt.match.PC) %>%
  mutate(Type = 'scrambled')

scramble.comparison.prompt <- rbind(prompt.real, prompt.scramble)

#shapiro.test(scramble.comparison.prompt$PC) # not normal

wilcox.test(subset(prompt.real, VMSgroup == "withVMS")$PC, 
            subset(prompt.scramble, VMSgroup == "withVMS")$PC, conf.int = T) # Significant, p<.001

cliff.delta(d = subset(prompt.real, VMSgroup == "withVMS")$PC, 
            f = subset(prompt.scramble, VMSgroup == "withVMS")$PC) # .90 large

wilcox.test(subset(prompt.real, VMSgroup == "noVMS")$PC, 
            subset(prompt.scramble, VMSgroup == "noVMS")$PC, conf.int = T) # Significant, p<.05

cliff.delta(d = subset(prompt.real, VMSgroup == "noVMS")$PC, 
            f = subset(prompt.scramble, VMSgroup == "noVMS")$PC) # .41 medium
```

```{r comparing inREP vs outREP consonants - nouns only}

#shapiro.test(subset(match.data.Prompt, VMSgroup=="withVMS")$PC) # normal 

# t.test(subset(match.data.Prompt_nouns.spread, VMSgroup=="withVMS")$inREP,
# subset(match.data.Prompt_nouns.spread, VMSgroup=="withVMS")$outREP, conf.int=T, paired = T) # p<.01

wilcox.test(subset(match.data.Prompt_nouns.spread, VMSgroup=="withVMS")$inREP, 
            subset(match.data.Prompt_nouns.spread, VMSgroup=="withVMS")$outREP, conf.int=T, paired = T) # p<.01

wilcoxonPairedR(x = subset(match.data.Prompt_nouns, VMSgroup == "withVMS")$PC,
                g = subset(match.data.Prompt_nouns, VMSgroup == "withVMS")$Type) # r = .58

```








